### 2024-09-24 paper/paper.qmd
initial commit

---
title: "Analyzing Toronto Island Ferry Ticket Counts"
subtitle: "A Statistical Exploration of Sales Trends"
author: 
  - Yingke He
thanks: "Code and data are available at: https://github.com/ohyykk/ferrySale.git"
date: 09-22-2024
date-format: long
abstract: "This study examines the ticket redemption and sales counts for the Toronto Island Ferry. We analyze time-based trends, relationships between sales and redemptions, and develop statistical models to predict ferry ticket usage. The results provide insights into usage patterns, highlight operational considerations, and suggest avenues for future research."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

# install.packages("here")

library(tidyverse)
library(opendatatoronto)
library(here)

```

# Introduction

The Toronto Island Ferry system has long been an essential transportation link between downtown Toronto and the Toronto Islands, serving both residents and tourists. With roots tracing back to the early days of the city’s development, the ferry system reflects both historical and contemporary mobility patterns in the region. Historical accounts provided by @varga1987toronto highlight the evolution of the Toronto Islands' natural environment and its significance to the area’s biodiversity and recreation. Additionally, the potential role of ferries in urban transit systems, as explored by @kopystynski1979potential, underscores their importance in enhancing urban mobility and connectivity.

This paper aims to analyze the ticket counts for the Toronto Island Ferry, exploring trends in ticket redemption and sales, and drawing connections between historical developments and current usage patterns. By integrating insights from both historical and modern perspectives, including contemporary studies on ferry commuting from @roseman2020precarious, this analysis seeks to provide a comprehensive understanding of the factors influencing ferry usage. The remainder of this paper is structured as follows: Section 2 describes the dataset used for analysis, Section 3 outlines the modeling approach, Section 4 presents the results, and Section 5 discusses the implications of the findings and potential directions for future research. We use @citeR and @rohan, and We use @sec-data to reference the data section.

# Data {#sec-data}

## Overview {#sec-data-overview}

The dataset used in this analysis is the 2024 installment of “Toronto Island Ferry Ticket Counts” obtained from OpenDataToronto (@varga1987toronto). This dataset contains information about the number of ferry tickets redeemed and sold for the Toronto Island Ferry system, which serves as a critical transportation link between downtown Toronto and the Toronto Islands.

The dataset is updated regularly and includes time-stamped entries for ticket redemptions and sales, allowing for a detailed analysis of ferry usage patterns throughout the year. The data is considered “open data” and can be utilized for various analytical purposes, provided that appropriate attribution is given..

The primary variables included in this analysis are:

Timestamp: The date and time when each ticket count was recorded.
Redemption.Count: The number of ferry tickets redeemed during the specified time period.
Sales.Count: The number of ferry tickets sold during the specified time period.

The data was processed and analyzed using the R programming language (@citeR). The tidyverse (@rohan) and opendatatoronto packages were employed to download, clean, and prepare the dataset for analysis. The resulting cleaned dataset was then utilized to generate the visualizations and statistical analyses presented in this paper.

## Results {#sec-data-results}

```{r}
#| echo: false
#| warning: false
#| message: false
#| results: "hide"

#### Loading the Cleaned Dataset ####
marriage_licence_packages <- search_packages("Toronto Island Ferry Ticket Counts")
marriage_licence_packages
# Retrieve a list of all resources (datasets) available within the found package
marriage_licence_resources <- marriage_licence_packages %>%
  list_package_resources()
# Display the list of resources available in the "Marriage Licence Statistics" package
marriage_licence_resources
marriage_licence_statistics <- marriage_licence_resources[2,] %>%
  get_resource()
dir.create(here::here("outputs"), showWarnings = FALSE)
dir.create(here::here("outputs/data"), showWarnings = FALSE)

# Save the retrieved dataset to a CSV file
write_csv(marriage_licence_statistics, here::here("outputs/data/a.csv"))
data=marriage_licence_statistics
data <- read_csv(here::here("outputs/data/a.csv"))

```

To gain an initial insight into the dataset, we plotted the trends in ticket redemption and sales counts over time. The line plot shown in @fig-linePlot provides a clear visual representation of these trends, serving as a starting point for deeper analysis of ferry usage patterns.

The dataset used in this analysis includes timestamped records of ticket redemptions and sales for the Toronto Island Ferry system, offering a time series perspective on ferry usage. The plot in @fig-linePlot specifically highlights how these counts fluctuate over time, reflecting the ebb and flow of passenger demand.

```{r}
#| label: fig-linePlot
#| fig-cap: Redemption and Sales Counts Over Time
#| echo: false

ggplot(data, aes(x = Timestamp)) +
  geom_line(aes(y = Redemption.Count, color = "Redemption Count")) +
  geom_line(aes(y = Sales.Count, color = "Sales Count")) +
  labs(title = "Redemption and Sales Counts Over Time",
       x = "Time",
       y = "Count") +
  scale_color_manual(name = "Legend", values = c("Redemption Count" = "blue", "Sales Count" = "red")) +
  theme_minimal()
```

As shown in @fig-linePlot, the x-axis denotes the timestamps of each record, while the y-axis represents the number of tickets processed. The blue line corresponds to the redemption counts, and the red line represents the sales counts.

The plot reveals significant variations in ferry usage over time. We observe several peaks in both redemption and sales counts, likely aligning with periods of high activity, such as weekends, holidays, or the peak tourist season during the summer months. These spikes suggest increased demand during favorable weather or event-driven occasions. Conversely, the noticeable dips in the graph may correspond to regular weekdays or off-season times, indicating lower ferry usage during these periods. This information is valuable for understanding the temporal dynamics of ferry operations and can inform better resource management and scheduling.
\newpage

The scatter plot in @fig-scatterPlot is designed to explore the relationship between ticket redemption and sales counts for the Toronto Island Ferry system. This visualization allows us to understand whether there is a correlation between the number of tickets sold and the number of tickets redeemed during the same time period.
```{r}
#| label: fig-scatterPlot
#| fig-cap: Scatter Plot of Redemption vs Sales Count
#| echo: false
#| warning: false
#| message: false

ggplot(data, aes(x = Redemption.Count, y = Sales.Count)) +
  geom_point(color = "purple", size = 3) +
  labs(title = "Scatter Plot of Redemption vs Sales Count",
       x = "Redemption Count",
       y = "Sales Count") +
  theme_minimal()
```

@fig-scatterPlot displays each data point as a purple dot, where the x-axis represents the count of tickets redeemed, and the y-axis denotes the count of tickets sold. This plot provides insight into the degree of alignment between ticket sales and actual usage (redemptions).

The scatter plot shows how ticket sales and redemptions relate to each other over time. If most points lie along a diagonal line, it indicates a strong positive correlation between the two variables, suggesting that a high number of tickets sold typically corresponds to a high number of tickets redeemed. This scenario could imply effective ticket utilization, where most purchased tickets are being used for travel.

\newpage

The box plot shown in @fig-boxPlot provides a summary of the distribution of ticket sales and redemption counts for the Toronto Island Ferry system. This visualization helps to identify the spread, central tendency, and any potential outliers in the data for both categories, offering a comparative view of the sales and redemption behaviors.
```{r}
#| label: fig-boxPlot
#| fig-cap: Box Plot of Sales and Redemption Counts
#| echo: false
#| warning: false
#| message: false

ggplot(data, aes(x = "Sales Count", y = Sales.Count, fill = "Sales Count")) +
  geom_boxplot(alpha = 0.5) +
  geom_boxplot(aes(x = "Redemption Count", y = Redemption.Count, fill = "Redemption Count"), alpha = 0.5) +
  scale_fill_manual(values = c("Sales Count" = "lightblue", "Redemption Count" = "lightgreen")) +
  labs(title = "Box Plot of Sales and Redemption Counts",
       x = "Category",
       y = "Count",
       fill = "Legend") +
  theme_minimal() +
  theme(legend.position = "top")
```
@fig-boxPlot illustrates two box plots side by side: one for sales counts (light blue) and the other for redemption counts (light green). Each box plot provides a visual summary of the data distribution, highlighting the median, quartiles, and any outliers for each category.

The box plots reveal key differences and similarities in the distribution of sales and redemption counts. The horizontal line inside each box represents the median count, providing a measure of central tendency. The top and bottom edges of the box denote the interquartile range (IQR), which captures the middle 50% of the data. The "whiskers" extending from the box show the range of values within 1.5 times the IQR from the first and third quartiles, while any points beyond these whiskers are considered outliers.

# Discussion {#sec-discussion}

## First discussion point {#sec-first-point}

In Section 2.2, the time series analysis of ticket sales and redemptions (Figure @fig-linePlot) revealed clear seasonal patterns, with pronounced peaks during the warmer months. This trend aligns with the historical and cultural significance of the Toronto Islands as a recreational destination, as documented by @varga1987toronto. The increased demand during these periods suggests that the ferry system plays a crucial role in supporting both tourism and local leisure activities. These seasonal fluctuations underscore the need for effective resource management and scheduling to accommodate surges in passenger volume during peak times.

## Second discussion point
The scatter plot analysis in Section 2.3 (Figure @fig-scatterPlot) demonstrated a generally positive correlation between ticket sales and redemptions, suggesting that most tickets purchased are used. However, discrepancies where sales exceed redemptions could be linked to bulk purchases, advanced ticket sales, or unredeemed tickets due to factors like adverse weather conditions. This observation is consistent with the analysis by @kopystynski1979potential, who highlighted the complexities of integrating ferry systems into urban transit networks. Understanding these discrepancies is critical for improving ticketing policies and service offerings to better match supply with demand.

## Third discussion point
The box plot analysis in Section 2.4 (Figure @fig-boxPlot) provided a comparative view of the variability in sales and redemption counts. The presence of outliers, particularly during periods of high sales but lower-than-expected redemptions, could indicate anomalies such as large-scale events or disruptions in service. Such variability highlights the need for a more robust system to handle unexpected fluctuations in demand, as well as a deeper understanding of passenger behavior. This finding resonates with the work of @roseman2020precarious, who explored the unpredictable nature of ferry commutes and the challenges faced by passengers in adapting to these changes.

## Weaknesses and next steps
Despite the insights gained, several limitations in the current analysis should be addressed. First, the dataset does not capture the full spectrum of ticket usage behavior, such as the impact of promotional offers or the use of multi-ride passes, which could lead to an overestimation or underestimation of actual ferry usage. Additionally, the analysis does not incorporate external factors such as weather conditions or special events that significantly influence ferry operations, as noted in prior studies like @mcilwraith1984transportation.

To address these limitations, future research should consider incorporating supplementary data sources, such as weather information, event schedules, and passenger feedback, to provide a more comprehensive view of the factors affecting ferry usage. Furthermore, a comparative study with other modes of transportation in Toronto could provide valuable insights into the role of the ferry system within the broader urban transit network. This would help identify areas for service improvement and inform strategies to enhance passenger experience and operational efficiency.

\newpage

\appendix

# Appendix {#sec-appendix}

## Dataset and Graph Sketches {#sec-appendix-sketches}

Sketches depicting both the desired dataset and the graphs generated in this analysis are available in the GitHub Repository.

## Data Cleaning {#sec-appendix-cleaning}

The data cleaning process involved filtering out some of the columns from the raw dataset and renaming some of the data entries for clarity and simplicity.

\newpage

# References {#sec-references}

### 2024-09-24 paper/paper.qmd
edit paper

@@
 thanks: "Code and data are available at: https://github.com/ohyykk/ferrySale.git"
-date: 09-22-2024
+date: 09-23-2024
 date-format: long

### 2024-09-25 outputs/paper/paper.qmd
changed folder summary

--- 2024-09-24 paper/paper.qmd
@@
 marriage_licence_packages <- search_packages("Toronto Island Ferry Ticket Counts")
-marriage_licence_packages
-# Retrieve a list of all resources (datasets) available within the found package
 marriage_licence_resources <- marriage_licence_packages %>%
   list_package_resources()
-# Display the list of resources available in the "Marriage Licence Statistics" package
-marriage_licence_resources
+
+# Retrieve the dataset
 marriage_licence_statistics <- marriage_licence_resources[2,] %>%
   get_resource()
-dir.create(here::here("outputs"), showWarnings = FALSE)
-dir.create(here::here("outputs/data"), showWarnings = FALSE)
 
-# Save the retrieved dataset to a CSV file
-write_csv(marriage_licence_statistics, here::here("outputs/data/a.csv"))
-data=marriage_licence_statistics
-data <- read_csv(here::here("outputs/data/a.csv"))
+# Create necessary directories
+dir.create(here::here("STA304/paper1/ferrySale/outputs"), showWarnings = FALSE, recursive = TRUE)
+
+# Save the dataset to a CSV file
+write_csv(marriage_licence_statistics, here::here("STA304/paper1/ferrySale/outputs/a.csv"))
+
+# Read the dataset back
+data <- read_csv(here::here("STA304/paper1/ferrySale/outputs/a.csv"))
 
@@
 
-ggplot(data, aes(x = Timestamp)) +
-  geom_line(aes(y = Redemption.Count, color = "Redemption Count")) +
-  geom_line(aes(y = Sales.Count, color = "Sales Count")) +
+# Plot Redemption and Sales Counts Over Time
+ggplot(data, aes(x = `Timestamp`)) +
+  geom_line(aes(y = `Redemption.Count`, color = "Redemption Count")) +
+  geom_line(aes(y = `Sales.Count`, color = "Sales Count")) +
   labs(title = "Redemption and Sales Counts Over Time",
@@
 This information is valuable for understanding the temporal dynamics of ferry operations and can inform better resource management and scheduling.
-\newpage
 
@@
 ggplot(data, aes(x = Redemption.Count, y = Sales.Count)) +
-  geom_point(color = "purple", size = 3) +
-  labs(title = "Scatter Plot of Redemption vs Sales Count",
-       x = "Redemption Count",
-       y = "Sales Count") +
-  theme_minimal()
+  geom_point(color = "purple", size = 3, alpha = 0.7) +  # Add transparency for overlapping points
+  labs(title = "Relationship Between Redemption and Sales Count",
+       subtitle = "Scatter plot with Redemption Count on X-axis and Sales Count on Y-axis",
+       x = "Redemption Count (units)",
+       y = "Sales Count (units)",
+       caption = "Source: Company Sales Data") +  # Add a caption for clarity
+  theme_minimal(base_size = 14) +  # Increase base font size for readability
+  theme(
+    plot.title = element_text(face = "bold", hjust = 0.5),  # Center and bold the title
+    plot.subtitle = element_text(hjust = 0.5),  # Center the subtitle
+    axis.title.x = element_text(face = "italic", margin = margin(t = 10)),  # Italicize X axis label and add margin
+    axis.title.y = element_text(face = "italic", margin = margin(r = 10)),  # Italicize Y axis label and add margin
+    panel.grid.major = element_line(color = "gray80", size = 0.5),  # Make grid lines more subtle
+    panel.grid.minor = element_blank(),  # Remove minor grid lines
+    plot.caption = element_text(hjust = 1, face = "italic", color = "gray40")  # Style caption
+  )
+
 ```

### 2024-09-26 outputs/paper/paper.qmd
Rearranged all files
removed unused folders and updated the new ones

--- 2024-09-24 paper/paper.qmd
@@
 We analyze time-based trends, relationships between sales and redemptions, and develop statistical models to predict ferry ticket usage.
-The results provide insights into usage patterns, highlight operational considerations, and suggest avenues for future research."
+The results provide insights into usage patterns, highlight operational considerations, and suggest avenues for future research.
+Our analysis reveals distinct seasonal peaks and troughs in ferry usage, correlating with tourism and weather conditions.
+Furthermore, the study identifies discrepancies between ticket sales and redemptions, suggesting potential improvements in ticketing strategies and resource allocation."
 format: pdf
 number-sections: true
+# The toc parameter used to create a table of contents was obtained from
+# Quarto documentation:
+# https://quarto.org/docs/reference/formats/pdf.html
+toc: true
 bibliography: references.bib
@@
 
-# install.packages("here")
+#install.packages("here")
 
@@
 marriage_licence_packages <- search_packages("Toronto Island Ferry Ticket Counts")
-marriage_licence_packages
-# Retrieve a list of all resources (datasets) available within the found package
 marriage_licence_resources <- marriage_licence_packages %>%
   list_package_resources()
-# Display the list of resources available in the "Marriage Licence Statistics" package
-marriage_licence_resources
+
+# Retrieve the dataset
 marriage_licence_statistics <- marriage_licence_resources[2,] %>%
   get_resource()
-dir.create(here::here("outputs"), showWarnings = FALSE)
-dir.create(here::here("outputs/data"), showWarnings = FALSE)
 
-# Save the retrieved dataset to a CSV file
-write_csv(marriage_licence_statistics, here::here("outputs/data/a.csv"))
-data=marriage_licence_statistics
-data <- read_csv(here::here("outputs/data/a.csv"))
+raw_data_path <- "C:/Users/yuqiu/ferrySale/ferrySale-main/inputs/data/raw_data.csv"
+cleaned_data_path <- "C:/Users/yuqiu/ferrySale/ferrySale-main/outputs/data/cleaned_data.csv"
+
+write_csv(marriage_licence_statistics, raw_data_path)
+
+raw_data <- read_csv(raw_data_path, show_col_types = FALSE)
+
+cleaned_data <- raw_data
+write_csv(cleaned_data, cleaned_data_path)
+data <- read_csv(cleaned_data_path, show_col_types = FALSE)
 
 ```
+
+## Sample Cleaned Data 
+The table @tbl-cleaned_data provides a detailed snapshot of the Toronto Island Ferry ticket transactions for a specific time period.
+It includes a unique ID for each transaction, the Time at which each transaction occurred, the number of Tickets Redeemed, the number of Tickets Sold, and the Date of the transaction.
+```{r}
+#| label: tbl-cleaned_data
+#| fig-cap: Sample Cleaned Data
+#| echo: false
+
+# Load necessary libraries
+library(knitr)
+library(readr)
+
+# Load the cleaned data from the CSV file (adjust the path as needed)
+data <- read_csv("C:/Users/yuqiu/ferrySale/ferrySale-main/outputs/data/cleaned_data.csv", show_col_types = FALSE)
+
+# Select a segment of the data, e.g., rows 51 to 60 for illustration
+segment_data <- data[51:60, ]
+
+# Modify the data for display - create new columns for ID, Time, and Date
+segment_data <- segment_data %>%
+  mutate(ID = row_number() + 50,                          # Create a new ID column starting from 51
+         Time = format(as.POSIXct(Timestamp), "%H:%M:%S"), # Extract Time from Timestamp
+         Date = format(as.POSIXct(Timestamp), "%Y-%m-%d")) # Extract Date from Timestamp
+
+# Select only the required columns for display in the table
+table_data <- segment_data[, c("ID", "Time", "Redemption.Count", "Sales.Count", "Date")]
+
+# Create the table using kable
+kable(table_data, col.names = c('ID', 'Time', 'Tickets Redeemed', 'Tickets Sold', 'Date'), 
+      align = "c", 
+      caption = "Sample of cleaned Ferry Ticket Counts Data")
+
+```
+
+This table shows the typical flucuations in ticket sales and redemptions throughout the day.
+For example, during late evening hours, we observe varying ticket redemption and sales patterns, which may correspond to changes in passenger demand as the ferry service approaches the end of its operational hours.
+
+## Redemption and Sales Counts Over Time
 
@@
 The dataset used in this analysis includes timestamped records of ticket redemptions and sales for the Toronto Island Ferry system, offering a time series perspective on ferry usage.
-The plot in @fig-linePlot specifically highlights how these counts fluctuate over time, reflecting the ebb and flow of passenger demand.
+The plot in @fig-linePlot specifically highlights how these counts fluctuate over time, reflecting the flow of passenger demand.
+The y-axis for redemption counts is displayed in blue, while the y-axis for sales counts is displayed in green, but inverted to better show their correlation.
+\newpage
 
@@
 
+# Calculate the range for both y-axes to have a symmetrical appearance
+y_max <- max(max(data$Sales.Count), max(data$Redemption.Count))
+
+# Create the plot
 ggplot(data, aes(x = Timestamp)) +
-  geom_line(aes(y = Redemption.Count, color = "Redemption Count")) +
-  geom_line(aes(y = Sales.Count, color = "Sales Count")) +
-  labs(title = "Redemption and Sales Counts Over Time",
-       x = "Time",
-       y = "Count") +
-  scale_color_manual(name = "Legend", values = c("Redemption Count" = "blue", "Sales Count" = "red")) +
-  theme_minimal()
+  # Line plot for Redemption Count
+  geom_line(aes(y = Redemption.Count), color = "blue", linewidth = 1) +
+  # Line plot for Sales Count inverted
+  geom_line(aes(y = -Sales.Count), color = "green", linewidth = 1) +
+  # Set y-axis limits symmetrically
+  scale_y_continuous(limits = c(-y_max, y_max),
+                     breaks = seq(-y_max, y_max, by = 1000),
+
+                     sec.axis = sec_axis(~ -., 
+                                         name = "Sales Count",
+                                         breaks = seq(y_max, -y_max, by = -1000))) +
+  labs(title = "Redemption and Sales Count Over Time",
+       y = "Redemption Count",
+       x = "Timestamp") +
+  theme_minimal() +
+  theme(axis.title.y.right = element_text(color = "green"),
+        axis.title.y.left = element_text(color = "blue"))
 ```
@@
 As shown in @fig-linePlot, the x-axis denotes the timestamps of each record, while the y-axis represents the number of tickets processed.
-The blue line corresponds to the redemption counts, and the red line represents the sales counts.
+The blue line corresponds to the redemption counts, and the green line represents the sales counts, plotted inversely to emphasize the contrast between the two trends.
+The symmetrical y-axis allows for a clearer comparison between peaks and troughs in both counts over time.
 
@@
 This information is valuable for understanding the temporal dynamics of ferry operations and can inform better resource management and scheduling.
+
 \newpage
+
+## Redemption vs Sales Count
 
@@
 
-ggplot(data, aes(x = Redemption.Count, y = Sales.Count)) +
-  geom_point(color = "purple", size = 3) +
-  labs(title = "Scatter Plot of Redemption vs Sales Count",
-       x = "Redemption Count",
-       y = "Sales Count") +
+ggplot(data, aes(x=Sales.Count, y=Redemption.Count)) +
+  geom_point(alpha=0.7) +
+  labs(title="Scatterplot: Redemption Count vs Sales Count",
+       x="Sales Count",
+       y="Redemption Count") +
+  xlim(0, max(data$Sales.Count) * 1.1) +  # Adding some space to the x-axis
+  ylim(0, max(data$Redemption.Count) * 1.1) +  # Adding some space to the y-axis
   theme_minimal()
+
 ```
 
-@fig-scatterPlot displays each data point as a purple dot, where the x-axis represents the count of tickets redeemed, and the y-axis denotes the count of tickets sold.
+@fig-scatterPlot displays each data point as a purple dot, where the y-axis represents the count of tickets redeemed, and the x-axis denotes the count of tickets sold.
 This plot provides insight into the degree of alignment between ticket sales and actual usage (redemptions).
@@
 \newpage
+
+## Side by Side Comparison
 
@@
 
-## First discussion point {#sec-first-point}
-
 In Section 2.2, the time series analysis of ticket sales and redemptions (Figure @fig-linePlot) revealed clear seasonal patterns, with pronounced peaks during the warmer months.
@@
 
-## Second discussion point
 The scatter plot analysis in Section 2.3 (Figure @fig-scatterPlot) demonstrated a generally positive correlation between ticket sales and redemptions, suggesting that most tickets purchased are used.
@@
 
-## Third discussion point
 The box plot analysis in Section 2.4 (Figure @fig-boxPlot) provided a comparative view of the variability in sales and redemption counts.

### 2024-09-26 outputs/paper/paper.qmd
Deleted unneeded code that writes to the csv files

@@
 
-raw_data_path <- "C:/Users/yuqiu/ferrySale/ferrySale-main/inputs/data/raw_data.csv"
-cleaned_data_path <- "C:/Users/yuqiu/ferrySale/ferrySale-main/outputs/data/cleaned_data.csv"
-
-write_csv(marriage_licence_statistics, raw_data_path)
-
-raw_data <- read_csv(raw_data_path, show_col_types = FALSE)
-
-cleaned_data <- raw_data
-write_csv(cleaned_data, cleaned_data_path)
-data <- read_csv(cleaned_data_path, show_col_types = FALSE)
+# Read the cleaned data
+data <- read_csv(file = here("outputs/data/cleaned_data.csv"), 
+                 show_col_types = FALSE)
 
@@
 # Load the cleaned data from the CSV file (adjust the path as needed)
-data <- read_csv("C:/Users/yuqiu/ferrySale/ferrySale-main/outputs/data/cleaned_data.csv", show_col_types = FALSE)
+data <- read_csv(file = here("outputs/data/cleaned_data.csv"), show_col_types = FALSE)
 

### 2024-09-26 outputs/paper/paper.qmd
added citations of used libraries and changed colors for the plots

@@
 library(here)
+library(ggplot2)
 
@@
 
-Timestamp: The date and time when each ticket count was recorded.
-Redemption.Count: The number of ferry tickets redeemed during the specified time period.
-Sales.Count: The number of ferry tickets sold during the specified time period.
+**Timestamp**: The date and time when each ticket count was recorded.  
+**Redemption.Count**: The number of ferry tickets redeemed during the specified time period.  
+**Sales.Count**: The number of ferry tickets sold during the specified time period.
 
 The data was processed and analyzed using the R programming language (@citeR).
-The tidyverse (@rohan) and opendatatoronto packages were employed to download, clean, and prepare the dataset for analysis.
-The resulting cleaned dataset was then utilized to generate the visualizations and statistical analyses presented in this paper.
+The tidyverse (@wickham2019tidyverse) and opendatatoronto packages were employed to download, clean, and prepare the dataset for analysis.
+The resulting cleaned dataset was then utilized to generate the visualizations and statistical analyses presented in this paper using a collection of libraries which included the following packages:
+- ggplot2(@wickham2016ggplot2)
+- knitr(@xie2023knitr)
+- readr(@wickham2023readr)
 
-## Results {#sec-data-results}
+## Results {#sec-data-resultsk}
 
@@
 
-## Sample Cleaned Data 
 The table @tbl-cleaned_data provides a detailed snapshot of the Toronto Island Ferry ticket transactions for a specific time period.
@@
 The y-axis for redemption counts is displayed in blue, while the y-axis for sales counts is displayed in green, but inverted to better show their correlation.
-\newpage
 
@@
   # Line plot for Redemption Count
-  geom_line(aes(y = Redemption.Count), color = "blue", linewidth = 1) +
-  # Line plot for Sales Count inverted
-  geom_line(aes(y = -Sales.Count), color = "green", linewidth = 1) +
-  # Set y-axis limits symmetrically
+  geom_line(aes(y = Redemption.Count), color = "#F7C4C1", linewidth = 1.2) +
+  # Line plot for Sales Count (inverted)
+  geom_line(aes(y = -Sales.Count), color = "#81B0B2", linewidth = 1.2) +
+
+  # Set y-axis limits symmetrically and apply the secondary axis
   scale_y_continuous(limits = c(-y_max, y_max),
                      breaks = seq(-y_max, y_max, by = 1000),
-
                      sec.axis = sec_axis(~ -., 
@@
                                          breaks = seq(y_max, -y_max, by = -1000))) +
-  labs(title = "Redemption and Sales Count Over Time",
-       y = "Redemption Count",
-       x = "Timestamp") +
+
+  # Add plot labels
+  labs(
+    title = "Redemption and Sales Count Over Time",
+    y = "Redemption Count",  # Y-axis for Redemption Count
+    x = "Timestamp"          # X-axis for Timestamp
+  ) +
+
+  # Apply minimal theme and customize axis titles
   theme_minimal() +
-  theme(axis.title.y.right = element_text(color = "green"),
-        axis.title.y.left = element_text(color = "blue"))
+  theme(
+    # Color the right-side axis title (Sales Count) to match the orange line
+    axis.title.y.right = element_text(color = "#F7C4C1", size = 12),
+    # Color the left-side axis title (Redemption Count) to match the pink line
+    axis.title.y.left = element_text(color = "#81B0B2", size = 12),
+    # Center the title and increase font size
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold")
+  )
 ```
@@
 
-\newpage
-
 ## Redemption vs Sales Count
@@
 
-ggplot(data, aes(x=Sales.Count, y=Redemption.Count)) +
-  geom_point(alpha=0.7) +
-  labs(title="Scatterplot: Redemption Count vs Sales Count",
-       x="Sales Count",
-       y="Redemption Count") +
+ggplot(data, aes(x = Sales.Count, y = Redemption.Count)) +
+  geom_point(alpha = 0.7, color = "#93445e") +
+  labs(
+    title = "Scatterplot: Redemption Count vs Sales Count",
+    x = "Sales Count",
+    y = "Redemption Count"
+  ) +
   xlim(0, max(data$Sales.Count) * 1.1) +  # Adding some space to the x-axis
   ylim(0, max(data$Redemption.Count) * 1.1) +  # Adding some space to the y-axis
-  theme_minimal()
-
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5)  # Center the title
+  )
 ```
@@
   geom_boxplot(aes(x = "Redemption Count", y = Redemption.Count, fill = "Redemption Count"), alpha = 0.5) +
-  scale_fill_manual(values = c("Sales Count" = "lightblue", "Redemption Count" = "lightgreen")) +
+  scale_fill_manual(values = c("Sales Count" = "lightgreen", "Redemption Count" = "lightblue")) +
   labs(title = "Box Plot of Sales and Redemption Counts",
@@
   theme_minimal() +
-  theme(legend.position = "top")
+  theme(
+    legend.position = "top",
+    plot.title = element_text(hjust = 0.5)  # Center the title
+  )
 ```

### 2024-09-26 outputs/paper/paper.qmd
made changes on text format, added bullet points

@@
 The resulting cleaned dataset was then utilized to generate the visualizations and statistical analyses presented in this paper using a collection of libraries which included the following packages:
-- ggplot2(@wickham2016ggplot2)
-- knitr(@xie2023knitr)
-- readr(@wickham2023readr)
+
+- ggplot2 [@wickham2016ggplot2]
+- knitr [@xie2023knitr]
+- readr [@wickham2023readr]
 

### 2024-09-26 outputs/paper/paper.tex
made changes on text format, added bullet points

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Analyzing Toronto Island Ferry Ticket Counts},
  pdfauthor={Yingke He},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Analyzing Toronto Island Ferry Ticket Counts\thanks{Code and data
are available at: https://github.com/ohyykk/ferrySale.git}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{A Statistical Exploration of Sales Trends}
\author{Yingke He}
\date{September 23, 2024}

\begin{document}
\maketitle
\begin{abstract}
This study examines the ticket redemption and sales counts for the
Toronto Island Ferry. We analyze time-based trends, relationships
between sales and redemptions, and develop statistical models to predict
ferry ticket usage. The results provide insights into usage patterns,
highlight operational considerations, and suggest avenues for future
research. Our analysis reveals distinct seasonal peaks and troughs in
ferry usage, correlating with tourism and weather conditions.
Furthermore, the study identifies discrepancies between ticket sales and
redemptions, suggesting potential improvements in ticketing strategies
and resource allocation.
\end{abstract}

\renewcommand*\contentsname{Table of contents}
{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{3}
\tableofcontents
}

\section{Introduction}\label{introduction}

The Toronto Island Ferry system has long been an essential
transportation link between downtown Toronto and the Toronto Islands,
serving both residents and tourists. With roots tracing back to the
early days of the city's development, the ferry system reflects both
historical and contemporary mobility patterns in the region. Historical
accounts provided by Varga (1987) highlight the evolution of the Toronto
Islands' natural environment and its significance to the area's
biodiversity and recreation. Additionally, the potential role of ferries
in urban transit systems, as explored by Kopystynski (1979), underscores
their importance in enhancing urban mobility and connectivity.

This paper aims to analyze the ticket counts for the Toronto Island
Ferry, exploring trends in ticket redemption and sales, and drawing
connections between historical developments and current usage patterns.
By integrating insights from both historical and modern perspectives,
including contemporary studies on ferry commuting from Roseman (2020),
this analysis seeks to provide a comprehensive understanding of the
factors influencing ferry usage. The remainder of this paper is
structured as follows: Section 2 describes the dataset used for
analysis, Section 3 outlines the modeling approach, Section 4 presents
the results, and Section 5 discusses the implications of the findings
and potential directions for future research. We use R Core Team (2023)
and Wickham et al. (2019a), and We use Section~\ref{sec-data} to
reference the data section.

\section{Data}\label{sec-data}

\subsection{Overview}\label{sec-data-overview}

The dataset used in this analysis is the 2024 installment of ``Toronto
Island Ferry Ticket Counts'' obtained from OpenDataToronto (Varga
(1987)). This dataset contains information about the number of ferry
tickets redeemed and sold for the Toronto Island Ferry system, which
serves as a critical transportation link between downtown Toronto and
the Toronto Islands.

The dataset is updated regularly and includes time-stamped entries for
ticket redemptions and sales, allowing for a detailed analysis of ferry
usage patterns throughout the year. The data is considered ``open data''
and can be utilized for various analytical purposes, provided that
appropriate attribution is given..

The primary variables included in this analysis are:

\textbf{Timestamp}: The date and time when each ticket count was
recorded.\\
\textbf{Redemption.Count}: The number of ferry tickets redeemed during
the specified time period.\\
\textbf{Sales.Count}: The number of ferry tickets sold during the
specified time period.

The data was processed and analyzed using the R programming language (R
Core Team (2023)). The tidyverse (Wickham et al. (2019b)) and
opendatatoronto packages were employed to download, clean, and prepare
the dataset for analysis. The resulting cleaned dataset was then
utilized to generate the visualizations and statistical analyses
presented in this paper using a collection of libraries which included
the following packages:

\begin{itemize}
\tightlist
\item
  ggplot2 (Wickham 2016)
\item
  knitr (Xie 2023)
\item
  readr (Wickham, Hester, and Bryan 2023)
\end{itemize}

\subsection{Results}\label{sec-data-resultsk}

The table Table~\ref{tbl-cleaned_data} provides a detailed snapshot of
the Toronto Island Ferry ticket transactions for a specific time period.
It includes a unique ID for each transaction, the Time at which each
transaction occurred, the number of Tickets Redeemed, the number of
Tickets Sold, and the Date of the transaction.

\begin{longtable}[]{@{}ccccc@{}}

\caption{\label{tbl-cleaned_data}Sample of cleaned Ferry Ticket Counts
Data}

\tabularnewline

\toprule\noalign{}
ID & Time & Tickets Redeemed & Tickets Sold & Date \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
51 & 23:30:00 & 5 & 4 & 2024-09-24 \\
52 & 23:15:00 & 1 & 3 & 2024-09-24 \\
53 & 23:00:00 & 2 & 0 & 2024-09-24 \\
54 & 22:45:00 & 7 & 3 & 2024-09-24 \\
55 & 22:30:00 & 11 & 0 & 2024-09-24 \\
56 & 22:15:00 & 2 & 1 & 2024-09-24 \\
57 & 22:00:00 & 5 & 6 & 2024-09-24 \\
58 & 21:45:00 & 2 & 3 & 2024-09-24 \\
59 & 21:30:00 & 14 & 41 & 2024-09-24 \\
60 & 21:15:00 & 3 & 1 & 2024-09-24 \\

\end{longtable}

Sample Cleaned Data

This table shows the typical flucuations in ticket sales and redemptions
throughout the day. For example, during late evening hours, we observe
varying ticket redemption and sales patterns, which may correspond to
changes in passenger demand as the ferry service approaches the end of
its operational hours.

\subsection{Redemption and Sales Counts Over
Time}\label{redemption-and-sales-counts-over-time}

To gain an initial insight into the dataset, we plotted the trends in
ticket redemption and sales counts over time. The line plot shown in
Figure~\ref{fig-linePlot} provides a clear visual representation of
these trends, serving as a starting point for deeper analysis of ferry
usage patterns.

The dataset used in this analysis includes timestamped records of ticket
redemptions and sales for the Toronto Island Ferry system, offering a
time series perspective on ferry usage. The plot in
Figure~\ref{fig-linePlot} specifically highlights how these counts
fluctuate over time, reflecting the flow of passenger demand. The y-axis
for redemption counts is displayed in blue, while the y-axis for sales
counts is displayed in green, but inverted to better show their
correlation.

\begin{figure}

\centering{

\includegraphics{paper_files/figure-pdf/fig-linePlot-1.pdf}

}

\caption{\label{fig-linePlot}Redemption and Sales Counts Over Time}

\end{figure}%

As shown in Figure~\ref{fig-linePlot}, the x-axis denotes the timestamps
of each record, while the y-axis represents the number of tickets
processed. The blue line corresponds to the redemption counts, and the
green line represents the sales counts, plotted inversely to emphasize
the contrast between the two trends. The symmetrical y-axis allows for a
clearer comparison between peaks and troughs in both counts over time.

The plot reveals significant variations in ferry usage over time. We
observe several peaks in both redemption and sales counts, likely
aligning with periods of high activity, such as weekends, holidays, or
the peak tourist season during the summer months. These spikes suggest
increased demand during favorable weather or event-driven occasions.
Conversely, the noticeable dips in the graph may correspond to regular
weekdays or off-season times, indicating lower ferry usage during these
periods. This information is valuable for understanding the temporal
dynamics of ferry operations and can inform better resource management
and scheduling.

\subsection{Redemption vs Sales Count}\label{redemption-vs-sales-count}

The scatter plot in Figure~\ref{fig-scatterPlot} is designed to explore
the relationship between ticket redemption and sales counts for the
Toronto Island Ferry system. This visualization allows us to understand
whether there is a correlation between the number of tickets sold and
the number of tickets redeemed during the same time period.

\begin{figure}

\centering{

\includegraphics{paper_files/figure-pdf/fig-scatterPlot-1.pdf}

}

\caption{\label{fig-scatterPlot}Scatter Plot of Redemption vs Sales
Count}

\end{figure}%

Figure~\ref{fig-scatterPlot} displays each data point as a purple dot,
where the y-axis represents the count of tickets redeemed, and the
x-axis denotes the count of tickets sold. This plot provides insight
into the degree of alignment between ticket sales and actual usage
(redemptions).

The scatter plot shows how ticket sales and redemptions relate to each
other over time. If most points lie along a diagonal line, it indicates
a strong positive correlation between the two variables, suggesting that
a high number of tickets sold typically corresponds to a high number of
tickets redeemed. This scenario could imply effective ticket
utilization, where most purchased tickets are being used for travel.

\newpage

\subsection{Side by Side Comparison}\label{side-by-side-comparison}

The box plot shown in Figure~\ref{fig-boxPlot} provides a summary of the
distribution of ticket sales and redemption counts for the Toronto
Island Ferry system. This visualization helps to identify the spread,
central tendency, and any potential outliers in the data for both
categories, offering a comparative view of the sales and redemption
behaviors.

\begin{figure}

\centering{

\includegraphics{paper_files/figure-pdf/fig-boxPlot-1.pdf}

}

\caption{\label{fig-boxPlot}Box Plot of Sales and Redemption Counts}

\end{figure}%

Figure~\ref{fig-boxPlot} illustrates two box plots side by side: one for
sales counts (light blue) and the other for redemption counts (light
green). Each box plot provides a visual summary of the data
distribution, highlighting the median, quartiles, and any outliers for
each category.

The box plots reveal key differences and similarities in the
distribution of sales and redemption counts. The horizontal line inside
each box represents the median count, providing a measure of central
tendency. The top and bottom edges of the box denote the interquartile
range (IQR), which captures the middle 50\% of the data. The
``whiskers'' extending from the box show the range of values within 1.5
times the IQR from the first and third quartiles, while any points
beyond these whiskers are considered outliers.

\section{Discussion}\label{sec-discussion}

In Section 2.2, the time series analysis of ticket sales and redemptions
(Figure Figure~\ref{fig-linePlot}) revealed clear seasonal patterns,
with pronounced peaks during the warmer months. This trend aligns with
the historical and cultural significance of the Toronto Islands as a
recreational destination, as documented by Varga (1987). The increased
demand during these periods suggests that the ferry system plays a
crucial role in supporting both tourism and local leisure activities.
These seasonal fluctuations underscore the need for effective resource
management and scheduling to accommodate surges in passenger volume
during peak times.

The scatter plot analysis in Section 2.3 (Figure
Figure~\ref{fig-scatterPlot}) demonstrated a generally positive
correlation between ticket sales and redemptions, suggesting that most
tickets purchased are used. However, discrepancies where sales exceed
redemptions could be linked to bulk purchases, advanced ticket sales, or
unredeemed tickets due to factors like adverse weather conditions. This
observation is consistent with the analysis by Kopystynski (1979), who
highlighted the complexities of integrating ferry systems into urban
transit networks. Understanding these discrepancies is critical for
improving ticketing policies and service offerings to better match
supply with demand.

The box plot analysis in Section 2.4 (Figure Figure~\ref{fig-boxPlot})
provided a comparative view of the variability in sales and redemption
counts. The presence of outliers, particularly during periods of high
sales but lower-than-expected redemptions, could indicate anomalies such
as large-scale events or disruptions in service. Such variability
highlights the need for a more robust system to handle unexpected
fluctuations in demand, as well as a deeper understanding of passenger
behavior. This finding resonates with the work of Roseman (2020), who
explored the unpredictable nature of ferry commutes and the challenges
faced by passengers in adapting to these changes.

\subsection{Weaknesses and next steps}\label{weaknesses-and-next-steps}

Despite the insights gained, several limitations in the current analysis
should be addressed. First, the dataset does not capture the full
spectrum of ticket usage behavior, such as the impact of promotional
offers or the use of multi-ride passes, which could lead to an
overestimation or underestimation of actual ferry usage. Additionally,
the analysis does not incorporate external factors such as weather
conditions or special events that significantly influence ferry
operations, as noted in prior studies like McIlwraith (1984).

To address these limitations, future research should consider
incorporating supplementary data sources, such as weather information,
event schedules, and passenger feedback, to provide a more comprehensive
view of the factors affecting ferry usage. Furthermore, a comparative
study with other modes of transportation in Toronto could provide
valuable insights into the role of the ferry system within the broader
urban transit network. This would help identify areas for service
improvement and inform strategies to enhance passenger experience and
operational efficiency.

\newpage

\appendix

\section{Appendix}\label{sec-appendix}

\subsection{Dataset and Graph Sketches}\label{sec-appendix-sketches}

Sketches depicting both the desired dataset and the graphs generated in
this analysis are available in the GitHub Repository.

\subsection{Data Cleaning}\label{sec-appendix-cleaning}

The data cleaning process involved filtering out some of the columns
from the raw dataset and renaming some of the data entries for clarity
and simplicity.

\newpage

\section*{References}\label{sec-references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-kopystynski1979potential}
Kopystynski, Adrian Danie. 1979. {``The Potential of Passenger Ferries
in an Urban Transit System.''} PhD thesis, University of British
Columbia.

\bibitem[\citeproctext]{ref-mcilwraith1984transportation}
McIlwraith, Thomas F. 1984. {``TRANSPORTATION IN OLD ONTARIO.''}
\emph{American Review of Canadian Studies} 14 (2): 177--92.

\bibitem[\citeproctext]{ref-citeR}
R Core Team. 2023. \emph{R: A Language and Environment for Statistical
Computing}. Vienna, Austria: R Foundation for Statistical Computing.
\url{https://www.R-project.org/}.

\bibitem[\citeproctext]{ref-roseman2020precarious}
Roseman, Sharon R. 2020. {``Precarious Mobility and Spectacle Dwelling
on a Ferry Commute.''} \emph{Applied Mobilities} 5 (1): 86--102.

\bibitem[\citeproctext]{ref-varga1987toronto}
Varga, Steve. 1987. \emph{Toronto Islands}. Toronto Field Naturalists.

\bibitem[\citeproctext]{ref-wickham2016ggplot2}
Wickham, Hadley. 2016. \emph{Ggplot2: Elegant Graphics for Data
Analysis}. Springer-Verlag New York.
\url{https://ggplot2.tidyverse.org}.

\bibitem[\citeproctext]{ref-rohan}
Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy
D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019a.
{``Welcome to the {tidyverse}.''} \emph{Journal of Open Source Software}
4 (43): 1686. \url{https://doi.org/10.21105/joss.01686}.

\bibitem[\citeproctext]{ref-wickham2019tidyverse}
Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy
D'Agostino McGowan, Romuald François, Garrett Grolemund, et al. 2019b.
{``Welcome to the Tidyverse.''} \emph{Journal of Open Source Software}.
\url{https://doi.org/10.21105/joss.01686}.

\bibitem[\citeproctext]{ref-wickham2023readr}
Wickham, Hadley, Jim Hester, and Jennifer Bryan. 2023. \emph{Readr: Read
Rectangular Text Data}. \url{https://CRAN.R-project.org/package=readr}.

\bibitem[\citeproctext]{ref-xie2023knitr}
Xie, Yihui. 2023. {``Knitr: A General-Purpose Package for Dynamic Report
Generation in r.''} \url{https://yihui.org/knitr/}.

\end{CSLReferences}

\end{document}

### 2024-09-26 outputs/paper/paper.qmd
removed unused files and edited the figcap of the table
deleted the titles for the three plots

@@
 #| label: tbl-cleaned_data
-#| fig-cap: Sample Cleaned Data
 #| echo: false
@@
   labs(
-    title = "Redemption and Sales Count Over Time",
     y = "Redemption Count",  # Y-axis for Redemption Count
@@
     # Center the title and increase font size
-    plot.title = element_text(hjust = 0.5, size = 16, face = "bold")
   )
@@
   labs(
-    title = "Scatterplot: Redemption Count vs Sales Count",
     x = "Sales Count",
@@
   scale_fill_manual(values = c("Sales Count" = "lightgreen", "Redemption Count" = "lightblue")) +
-  labs(title = "Box Plot of Sales and Redemption Counts",
+  labs(
        x = "Category",

### 2024-09-27 scripts/00-simulate_data.R
finished simulating the basic raw data

@@ -1,0 +1,40 @@
+#### Preamble ####
+# Purpose: Simulates a dataset representing Toronto's Island Ferry Ticket Counts
+# Author: Yingke He
+# Date: 26 September 2024
+# Contact: kiki.he@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: NA
+# Any other information needed: NA
+
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(dplyr)
+
+
+
+#### Simulate data ####
+# Set seed for reproducibility
+set.seed(123)
+
+# Number of rows to simulate
+n <- 100
+
+# Simulating the data
+simulated_data <- data.frame(
+  X_id = 1:n,
+  Timestamp = seq.POSIXt(from = as.POSIXct("2024-09-25 16:45:00"),
+                         by = "15 mins", length.out = n),
+  Redemption.Count = round(rnorm(n, mean = 35, sd = 10)),  # Mean and SD based on observed values
+  Sales.Count = round(rnorm(n, mean = 32, sd = 10))        # Mean and SD based on observed values
+)
+
+# Ensure that counts are non-negative
+simulated_data <- simulated_data %>%
+  mutate(Redemption.Count = ifelse(Redemption.Count < 0, 0, Redemption.Count),
+         Sales.Count = ifelse(Sales.Count < 0, 0, Sales.Count))
+
+# Print the first few rows of the simulated data for checking purposes
+head(simulated_data)

### 2024-09-27 scripts/01-download_data.R
updated missing package in download data file

@@ -1,0 +1,40 @@
+#### Preamble ####
+# Purpose: Downloads and saves the data from the FerrySale dataset package
+# Author: Yingke He
+# Date: 26 September 2024
+# Contact: kiki.he@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: NA
+# Any other information needed: NA
+
+#### Workspace setup ####
+library(opendatatoronto)
+library(dplyr)
+library(readr)
+
+# Get package using the search term "Toronto Island Ferry Ticket Counts"
+ferry_ticket_packages <- search_packages("Toronto Island Ferry Ticket Counts")
+
+# Get all resources for this package (using the package ID for the "Toronto Island Ferry Ticket Counts")
+resources <- list_package_resources(ferry_ticket_packages$id[1])
+
+# Identify the datastore resources
+datastore_resources <- resources %>%
+  filter(tolower(format) == "datastore")
+
+# Load the first datastore resource as a sample
+data <- datastore_resources %>%
+  filter(row_number() == 1) %>%
+  get_resource()
+
+# Save the loaded data
+raw_data_path <- "/Users/kiki/Desktop/STA304/paper1/ferrySale/scripts/raw_data.csv"
+
+# Write the raw data to a CSV file
+write_csv(data, raw_data_path)
+
+# Read the saved raw data
+raw_data <- read_csv(raw_data_path, show_col_types = FALSE)
+
+#### Save data ####
+write_csv(raw_data, "data/raw_data/raw_data.csv")

### 2024-09-27 scripts/02-data_cleaning.R
updated saving directory and file names

@@ -1,0 +1,30 @@
+#### Preamble ####
+# Purpose: Cleans the FerrySale data obtained
+# Author: Yingke He
+# Date: 26 September 2024
+# Contact: kiki.he@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: NA
+# Any other information needed: NA
+
+#### Workspace Set Up ####
+
+library(tidyverse)
+library(opendatatoronto)
+
+#### Cleaning the data set ####
+# read file first
+raw_data <- read_csv("data/raw_data/raw_data.csv")
+
+# Data Cleaning Steps # 1. Remove rows with missing values 
+raw_data <- raw_data %>% drop_na() 
+# 2. Convert Timestamp to proper date-time format 
+raw_data <- raw_data %>% mutate(Timestamp = as.POSIXct(Timestamp, format = "%Y-%m-%d %H:%M:%S")) 
+# 3. Remove duplicates 
+raw_data <- raw_data %>% distinct() 
+# 4. Filter out unrealistic values (example: negative sales or redemption counts) 
+raw_data <- raw_data %>% filter(Redemption.Count >= 0, Sales.Count >= 0)
+
+cleaned_data <- raw_data
+write_csv(cleaned_data, cleaned_data_path)
+data <- read_csv(cleaned_data_path, show_col_types = FALSE)

### 2024-09-27 scripts/03-test_data.R
Added more testing functions to test the data
Checked that all timestamps fall within the specified range

@@ -1,0 +1,78 @@
+#### Preamble ####
+# Purpose: Tests the integrity of raw and processed FerrySale data
+# Author: Yingke He
+# Date: 26 September 2024
+# Contact: kiki.he@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: NA
+# Any other information needed: NA
+
+#### Workspace setup ####
+library(testthat)
+library(readr)
+library(lubridate)
+
+#### Load data ####
+# Replace the paths below with your actual data paths
+ferry_data <- read_csv("data/raw_data/unedited_data.csv")
+
+#### Test data ####
+
+# Test Suite for Ferry Data
+test_that("Check column types in ferry_data dataset", {
+  
+  # Test that 'Timestamp' column is in DateTime format
+  expect_true(class(ferry_data$Timestamp) == "character", 
+              info = "'Timestamp' column should initially be in character format")
+  
+  # Convert 'Timestamp' to Date-Time and validate
+  ferry_data$Timestamp <- ymd_hms(ferry_data$Timestamp)
+  expect_true(class(ferry_data$Timestamp) == "POSIXct", 
+              info = "'Timestamp' column should be in POSIXct (DateTime) format after conversion")
+  
+  # Test that 'Redemption.Count' and 'Sales.Count' columns are numeric
+  expect_true(is.numeric(ferry_data$Redemption.Count), 
+              info = "'Redemption.Count' column should be numeric")
+  expect_true(is.numeric(ferry_data$Sales.Count), 
+              info = "'Sales.Count' column should be numeric")
+})
+
+# Check that 'X_id', 'Redemption.Count', and 'Sales.Count' columns do not have any missing values
+test_that("Check for missing values in key columns", {
+  expect_false(any(is.na(ferry_data$X_id)), 
+               info = "'X_id' column should not have missing values")
+  expect_false(any(is.na(ferry_data$Redemption.Count)), 
+               info = "'Redemption.Count' column should not have missing values")
+  expect_false(any(is.na(ferry_data$Sales.Count)), 
+               info = "'Sales.Count' column should not have missing values")
+})
+
+# Test to check that all 'Redemption.Count' and 'Sales.Count' values are non-negative
+test_that("Check that counts are non-negative", {
+  expect_true(all(ferry_data$Redemption.Count >= 0), 
+              info = "'Redemption.Count' should be non-negative")
+  expect_true(all(ferry_data$Sales.Count >= 0), 
[…]

### 2024-09-27 outputs/paper/paper.qmd
edit intro section, add a summary statistics table to the data section

@@
 ---
-title: "Analyzing Toronto Island Ferry Ticket Counts"
-subtitle: "A Statistical Exploration of Sales Trends"
+title: "Uncovering Ferry Traffic Patterns: A Data-Driven Study of Toronto Island Ticket Sales and Redemptions"
 author: 
@@
 date-format: long
-abstract: "This study examines the ticket redemption and sales counts for the Toronto Island Ferry.
-We analyze time-based trends, relationships between sales and redemptions, and develop statistical models to predict ferry ticket usage.
+abstract: "This study examines the ticket redemption and sales counts for the Toronto Island Ferry, analyzing time-based trends and the relationships between sales and redemptions.
+Statistical models are developed to predict ferry ticket usage.
 The results provide insights into usage patterns, highlight operational considerations, and suggest avenues for future research.
-Our analysis reveals distinct seasonal peaks and troughs in ferry usage, correlating with tourism and weather conditions.
+The analysis reveals distinct seasonal peaks and troughs in ferry usage, correlating with tourism and weather conditions.
 Furthermore, the study identifies discrepancies between ticket sales and redemptions, suggesting potential improvements in ticketing strategies and resource allocation."
@@
 # Introduction
-
 The Toronto Island Ferry system has long been an essential transportation link between downtown Toronto and the Toronto Islands, serving both residents and tourists.
@@
 Historical accounts provided by @varga1987toronto highlight the evolution of the Toronto Islands' natural environment and its significance to the area’s biodiversity and recreation.
-Additionally, the potential role of ferries in urban transit systems, as explored by @kopystynski1979potential, underscores their importance in enhancing urban mobility and connectivity.
+Additionally, the potential role of ferries in urban transit systems, as explained by @kopystynski1979potential, shows their importance in enhancing urban mobility and connectivity.
 
-This paper aims to analyze the ticket counts for the Toronto Island Ferry, exploring trends in ticket redemption and sales, and drawing connections between historical developments and current usage patterns.
-By integrating insights from both historical and modern perspectives, including contemporary studies on ferry commuting from @roseman2020precarious, this analysis seeks to provide a comprehensive understanding of the factors influencing ferry usage.
-The remainder of this paper is structured as follows: Section 2 describes the dataset used for analysis, Section 3 outlines the modeling approach, Section 4 presents the results, and Section 5 discusses the implications of the findings and potential directions for future research.
-We use @citeR and @rohan, and We use @sec-data to reference the data section.
+With the ticket counts for the Toronto Island Ferry and the trends in ticket redemption and sales, we can find the connections between historical developments and current usage patterns.
+By combining both historical and modern perspectives and including contemporary studies on ferry commuting from @roseman2020precarious, the analysis provides a comprehensive understanding of the factors influencing ferry usage.
+Statistical models are used to analyze ferry ticket demand while showing the seasonal fluctuations and operational challenges.
+
+Despite the longstanding role of the Toronto Island Ferry system in connecting downtown Toronto with the Toronto Islands, there is limited research on the detailed patterns of ferry usage and the efficiency of the service.
+While the ferry system serves both residents and tourists, the lack of data-driven analysis makes it harder to optimize ferry operations, manage peak demand periods, and improve ticketing strategies.
+It is important to fill this gap because of the ferry’s dual function of accommodating daily commuters alongside a significant influx of seasonal visitors, which introduces variability in service demand.
+Without a comprehensive understanding of the underlying trends in ticket sales and redemptions, it becomes challenging to address inefficiencies or anticipate future operational challenges.
+We will address this problem by providing a detailed analysis of ticketing patterns and offering insights into how the ferry system can better meet passenger needs through data-informed decision-making.
+
+In addition to the expected peaks during high tourism periods, this study investigates inconsistencies between tickets sold and redeemed, offering insights into potential service inefficiencies.
+The dual role of the Toronto Island Ferry system in accommodating both daily commuters and a large number of seasonal visitors makes it an important case study in urban transportation planning.
+
+The remainder of this paper is structured as follows: Section 2 introduces the dataset used for the analysis, explaining the key variables and presenting key data analysis.
+Section 3 discusses the implications of the findings and potential directions for future research.
+Additional information is presented in the appendix at section A.
+@citeR and @rohan is utilized for the coding proposes of this paper, and the data section is referenced using @sec-data.
 
@@
 The data was processed and analyzed using the R programming language (@citeR).
-The tidyverse (@wickham2019tidyverse) and opendatatoronto packages were employed to download, clean, and prepare the dataset for analysis.
+The tidyverse (@tidyverse2019) and opendatatoronto packages were employed to download, clean, and prepare the dataset for analysis.
 The resulting cleaned dataset was then utilized to generate the visualizations and statistical analyses presented in this paper using a collection of libraries which included the following packages:
@@
 - readr [@wickham2023readr]
+
+## Measurements
+
+This report utilizes the Toronto Island Ferry Ticket Counts dataset published by the City of Toronto on the Open Data Portal(@toronto_ferry_ticket_counts_2024).
+Specifically, time-stamped records of both ticket sales and redemptions were recoreded.
+Every time a ticket is sold or redeemed, it represents a physical action by a ferry passenger in Toronto traveling to the Toronto Island.
+Actions are then recorded by the ticketing systems, which automatically records the transaction timestamp, along with the number of tickets sold and redeemed at that specific time.
+
+This process converts individual passenger transactions into structured dataset entries, with timestamps to enable tracking over time.
+The data captures real-world trends such as higher ferry activity during peak tourism seasons or weekends, providing useful insights into ferry usage patterns.
+It enables further research to analyze passenger demand and operational trends, helping to make more effective ferry management decisions.
 
@@
       align = "c", 
-      caption = "Sample of cleaned Ferry Ticket Counts Data")
+      caption = "Table 1: Sample of cleaned Ferry Ticket Counts Data")
 
@@
 
-## Redemption and Sales Counts Over Time
+```{r}
+#| label: tbl-summary_stats
+#| echo: false
+#| results: 'asis'
+
+# Calculate summary statistics
+summary_stats <- data %>% 
+  summarise(
+    `Avg Redeemed` = mean(Redemption.Count, na.rm = TRUE), 
+    `Avg Sold` = mean(Sales.Count, na.rm = TRUE), 
+    `Max Redeemed` = max(Redemption.Count, na.rm = TRUE), 
+    `Max Sold` = max(Sales.Count, na.rm = TRUE), 
+    Count = n()
+  )
+
+# Create the table using kable
+knitr::kable(summary_stats, col.names = c('Avg Redeemed', 'Avg Sold', 'Max Redeemed', 'Max Sold', 'Count'), 
+      align = "c", caption = "Table 2: Summary Statistics of Ticket Counts")
+
+```
+The summary statistics, shown in Table 2, provide a clear understanding of ferry ticket trends.
+The high volume of tickets sold and redeemed indicates strong demand for the ferry service, with the close alignment between sales and redemptions suggesting high customer engagement.
+On average, 46.36 tickets were sold every 15 minutes during working hours, and 48.45 tickets were redeemed, highlighting that most customers have used their tickets.
+The small gap between sales and redemptions implies a few no-shows or unused tickets.
+The maximum values of 7,216 for sales and 7,229 for redemptions demonstrate the ferry’s capacity to handle high passenger volumes, typically operating at full capacity on busy days.
+
+### Temporal Trends in Redemption and Sales Counts
 
@@
 
-## Redemption vs Sales Count
+### Comparative Analysis of Ticket Sales and Redemptions
 
@@
 
-\newpage
-
-## Side by Side Comparison
+### Visual Comparison of Ticket Sales and Redemption Distribution
 

### 2024-09-27 outputs/paper/paper.qmd
made changes to the abstract

@@
 Statistical models are developed to predict ferry ticket usage.
-The results provide insights into usage patterns, highlight operational considerations, and suggest avenues for future research.
-The analysis reveals distinct seasonal peaks and troughs in ferry usage, correlating with tourism and weather conditions.
-Furthermore, the study identifies discrepancies between ticket sales and redemptions, suggesting potential improvements in ticketing strategies and resource allocation."
+The results uncover seasonal trends, operational inefficiencies, and discrepancies between ticket sales and redemptions, highlighting the need for improved resource management and service adjustments to better align with fluctuating passenger demand.
+These findings offer insights into usage patterns, operational considerations, and suggest avenues for future research, particularly in optimizing ticketing strategies and resource allocation."
 format: pdf
@@
 
-The remainder of this paper is structured as follows: Section 2 introduces the dataset used for the analysis, explaining the key variables and presenting key data analysis.
-Section 3 discusses the implications of the findings and potential directions for future research.
-Additional information is presented in the appendix at section A.
+The remainder of this paper is structured as follows: @sec-data introduces the dataset used for the analysis, explaining the key variables and presenting key data analysis.
+@sec-discussion discusses the implications of the findings and potential directions for future research.
+Additional information is presented in the appendix at @sec-appendix.
 @citeR and @rohan is utilized for the coding proposes of this paper, and the data section is referenced using @sec-data.
@@
 
-The table @tbl-cleaned_data provides a detailed snapshot of the Toronto Island Ferry ticket transactions for a specific time period.
+@tbl-cleaned_data provides a detailed snapshot of the Toronto Island Ferry ticket transactions for a specific time period.
 It includes a unique ID for each transaction, the Time at which each transaction occurred, the number of Tickets Redeemed, the number of Tickets Sold, and the Date of the transaction.
@@
       align = "c", 
-      caption = "Table 1: Sample of cleaned Ferry Ticket Counts Data")
+      caption = "Sample of cleaned Ferry Ticket Counts Data")
 
@@
 
-This table shows the typical flucuations in ticket sales and redemptions throughout the day.
+The above table shows the typical flucuations in ticket sales and redemptions throughout the day.
 For example, during late evening hours, we observe varying ticket redemption and sales patterns, which may correspond to changes in passenger demand as the ferry service approaches the end of its operational hours.
@@
 knitr::kable(summary_stats, col.names = c('Avg Redeemed', 'Avg Sold', 'Max Redeemed', 'Max Sold', 'Count'), 
-      align = "c", caption = "Table 2: Summary Statistics of Ticket Counts")
+      align = "c", caption = "Summary Statistics of Ticket Counts")
 
 ```
-The summary statistics, shown in Table 2, provide a clear understanding of ferry ticket trends.
+The summary statistics, shown in @tbl-summary_stats, provide a clear understanding of ferry ticket trends.
 The high volume of tickets sold and redeemed indicates strong demand for the ferry service, with the close alignment between sales and redemptions suggesting high customer engagement.
@@
 
-The scatter plot analysis in Section 2.3 (Figure @fig-scatterPlot) demonstrated a generally positive correlation between ticket sales and redemptions, suggesting that most tickets purchased are used.
+The scatter plot analysis in Section 2.3 (@fig-scatterPlot) demonstrated a generally positive correlation between ticket sales and redemptions, suggesting that most tickets purchased are used.
 However, discrepancies where sales exceed redemptions could be linked to bulk purchases, advanced ticket sales, or unredeemed tickets due to factors like adverse weather conditions.

### 2024-09-27 outputs/paper/paper.qmd
Plots added to section2.3.1
updated peak day summary table, Redemption and Sales Counts Over Time in days, months and a year

@@
 library(ggplot2)
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, error = FALSE)
 
@@
 
+suppressMessages({
+  suppressWarnings({
     #### Loading the Cleaned Dataset ####
@@
                      show_col_types = FALSE)
+  })
+})
 
@@
 
-### Temporal Trends in Redemption and Sales Counts
+### Temporal Trends in Redemption and Sales Counts {#sec-data-temporal}
 
 To gain an initial insight into the dataset, we plotted the trends in ticket redemption and sales counts over time.
-The line plot shown in @fig-linePlot provides a clear visual representation of these trends, serving as a starting point for deeper analysis of ferry usage patterns.
+The line plot shown in @fig-linePlot1, @fig-linePlot2, @fig-linePlot3 and @fig-linePlot4 provides a clear visual representation of these trends, serving as a starting point for deeper analysis of ferry usage patterns.
 
-The dataset used in this analysis includes timestamped records of ticket redemptions and sales for the Toronto Island Ferry system, offering a time series perspective on ferry usage.
-The plot in @fig-linePlot specifically highlights how these counts fluctuate over time, reflecting the flow of passenger demand.
-The y-axis for redemption counts is displayed in blue, while the y-axis for sales counts is displayed in green, but inverted to better show their correlation.
+The peak day analysis table below highlights the highest ferry ticket sales, often aligning with public holidays or major events, such as Aug 1 2016, and July 1 2018.
+On September 15, 2020, ferry sales saw a significant drop, with only 5,003 tickets sold.
+This decline could be due to the impact of COVID-19(@lee2021island).
 
 ```{r}
-#| label: fig-linePlot
-#| fig-cap: Redemption and Sales Counts Over Time
+
+#| label: tbl-peak_day
+#| echo: false       # Prevents code from being shown
+#| warning: false    # Suppresses warnings
+#| message: false    # Suppresses messages
+#| error: false      # Suppresses errors
+#| results: 'asis'   # Ensures the table result is shown without the code
+
+# Load necessary libraries
+library(dplyr)
+library(lubridate)
+library(knitr)
+
+# Step 1: Check if the Timestamp column is properly parsed
+data <- data %>%
+  mutate(Timestamp = ymd_hms(Timestamp, quiet = TRUE)) %>%  # Quiet mode for parsing errors
+  filter(!is.na(Timestamp))  # Remove rows where Timestamp conversion failed
+
+# Step 2: Group data by Year, find the day with maximum Sales for each year
+peak_day_per_year_table <- data %>%
+  mutate(Year = year(Timestamp)) %>%  # Extract Year from Timestamp
+  group_by(Year, Date = as.Date(Timestamp)) %>%  # Group by Year and Date
+  summarise(Total_Sales = sum(Sales.Count, na.rm = TRUE), .groups = 'drop') %>%  # Summarize total sales for each day
+  filter(Total_Sales > 0) %>%  # Ensure there are sales
+  group_by(Year) %>%
+  slice_max(order_by = Total_Sales, n = 1, with_ties = FALSE) %>%  # Find the day with maximum sales for each year
+  select(Year, Peak_Day = Date, Total_Sales)  # Select relevant columns
+
+# Step 3: Display the table using knitr::kable
+kable(peak_day_per_year_table, col.names = c("Year", "Peak Day", "Total Sales"), 
+      caption = "Table: Peak Day for Each Year Based on Total Sales")
+
+```
+
+The dataset used in this analysis provides timestamped records of ticket redemptions and sales for the Toronto Island Ferry system, allowing for a detailed time series analysis of ferry usage patterns.
+As illustrated in @fig-linePlot1, ferry usage fluctuates significantly throughout the year, with peak activity consistently observed between July and September.
+This pattern corresponds to the increased passenger demand during the summer holiday season and major events.
+In 2020, ticket demand saw a sharp decline, likely due to the impact of the COVID-19 pandemic(@urbanyi2020maritime), but it showed signs of gradual recovery starting in 2021 as restrictions eased and tourism began to rebound in Toronto(@travel_gov_2024).
+By identifying these peak traffic periods in advance, ferry operators can implement targeted operational strategies, such as scheduling additional ferry trips or introducing crowd control measures to minimize the risk of overcrowding or service disruptions.
+Identifing these high-demand periods can ensure smoother ferry operations and enhance the overall passenger experience.
+
+```{r}
+#| label: fig-linePlot1
+#| fig-cap: Redemption and Sales Counts Over Time in days
 #| echo: false
+#| warning: false
+#| message: false
+#| fig-width: 8
+#| fig-height: 4
+#| 
+# Load necessary libraries
+library(ggplot2)
+library(dplyr)
+library(lubridate)
+
+# Assuming 'data' has columns: 'Timestamp' and 'Sales.Count'
+# Convert Timestamp column to proper datetime format and filter out NA values
+data <- data %>%
+  mutate(
+    Timestamp = ymd_hms(Timestamp),   # Convert to datetime format
+    Year = year(Timestamp),           # Extract Year
+    Month = month(Timestamp, label = TRUE, abbr = TRUE),  # Extract Month (abbreviated)
+    Day = day(Timestamp)              # Extract Day
+  ) %>%
+  filter(!is.na(Year) & !is.na(Month))  # Remove rows with NA in Year or Month
+
+# Group the data by Year, Month, and Day, and calculate the total sales per day for each year
+daily_data <- data %>%
+  group_by(Year, Month, Day) %>%
+  summarise(
+    Total_Sales = sum(Sales.Count, na.rm = TRUE),
+    .groups = 'drop'  # Ungroup after summarising
+  )
+
+# Find the peak sales day for each year
+peak_day_data <- daily_data %>%
+  group_by(Year) %>%
+  filter(Total_Sales == max(Total_Sales)) %>%
+  ungroup()
+
+# Step 4: Define colors for each year
+year_colors <- c("2015" = "#d07f7a", "2016" = "#377EB8", "2017" = "#7f9f6f", 
+                 "2018" = "#984EA3", "2019" = "#FC9272", "2020" = "#EC7014", 
+                 "2021" = "#A65628", "2022" = "#C51B7D", "2023" = "#FFD92F")
+
+# Create the plot
+ggplot(peak_day_data, aes(x = Month, y = Total_Sales, group = Year, color = as.factor(Year))) +  
+  geom_line(size = 1.2) +   # Line plot for each year’s peak day
+  geom_point(size = 3) +    # Add points for peak days
+  # Set y-axis label and title
+  labs(
+    x = "Month",
+    y = "Total Sales Count",
+    color = "Year"
+  ) +
+
+  # Manually specify colors for each year
+  scale_color_manual(values = year_colors) +
+
+  # Apply minimal theme and customize axis titles
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16),
+    axis.title.x = element_text(size = 12),
+    axis.title.y = element_text(size = 12),
+    legend.title = element_text(size = 12),
+    legend.text = element_text(size = 10),
+    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels for better readability
+  )
+
+```
+
+@fig-linePlot2 below shows the Redemption and Sales Counts in months over years 加一段描述plot2
+
+\newpage
+
+```{r}
+#| label: fig-linePlot2
+#| fig-cap: Redemption and Sales Counts Over Time in months
+#| warnings: false
+#| message: false
+#| echo: false
+#| fig-pos: "H"  #fix the plot in current place when render
+#| fig-width: 8
+#| fig-height: 4
+#| 
+# Load necessary libraries
+library(ggplot2)
+library(dplyr)
+library(lubridate)
+
+# Convert Timestamp column to proper datetime format and filter out NA values
+data <- data %>%
+  mutate(
+    Timestamp = ymd_hms(Timestamp),     # Convert to datetime format
+    Year = year(Timestamp),             # Extract Year
+    Month = month(Timestamp, label = TRUE, abbr = TRUE)  # Extract Month (abbreviated)
+  ) %>%
+  filter(!is.na(Year) & !is.na(Month))  # Remove rows with NA in Year or Month
+
+# Step 2: Group the data by Year and Month and calculate total sales and redemptions per month for each year
+monthly_data <- data %>%
+  group_by(Year, Month) %>%
+  summarise(
+    Sales.Count = sum(Sales.Count, na.rm = TRUE),
+    Redemption.Count = sum(Redemption.Count, na.rm = TRUE),
+    .groups = 'drop'  # Ungroup after summarising
+  )
+
+# Step 3: Determine the maximum value for symmetrical scaling
+max_count <- max(max(monthly_data$Sales.Count), max(monthly_data$Redemption.Count))
+
+# Step 4: Define colors for each year
+year_colors <- c("2015" = "#d07f7a", "2016" = "#377EB8", "2017" = "#7f9f6f", 
+                 "2018" = "#984EA3", "2019" = "#FC9272", "2020" = "#EC7014", 
+                 "2021" = "#A65628", "2022" = "#C51B7D", "2023" = "#FFD92F")
+
+# Step 5: Create the plot with inverted Redemption Count for bottom half visualization
+ggplot(monthly_data, aes(x = Month)) +  
+  # Line plot for Redemption Count (inverted)
+  geom_line(aes(y = -Redemption.Count, color = as.factor(Year), group = Year), linewidth = 1.2) +
+  # Line plot for Sales Count (normal)
+  geom_line(aes(y = Sales.Count, color = as.factor(Year), group = Year), linetype = "dashed", linewidth = 1.2) +
+
+  # Set symmetrical y-axis limits for both Sales and Redemption
+  scale_y_continuous(limits = c(-max_count, max_count),
+                     breaks = seq(-max_count, max_count, by = 100000),
+                     sec.axis = sec_axis(~ -., name = "Sales Count (dashed lines)")) +
+
+  # Add plot labels
+  labs(
+    y = "Redemption Count (straight lines)",  # Y-axis for Redemption Count (inverted)
+    x = "Month",             # X-axis for Month
+    color = "Year"
+  ) +
+
+  # Manually specify colors for each year
+  scale_color_manual(values = year_colors) +
+
+  # Apply minimal theme and customize axis titles
+  theme_minimal() +
+  theme(
+    axis.title.y.right = element_text(color = "black", size = 12, margin = margin(r = 10)),
+    axis.title.y.left = element_text(color = "black", size = 12, margin = margin(l = 10)),
+    plot.title = element_text(hjust = 0.5, size = 16),
+    legend.title = element_text(size = 12),
+    legend.text = element_text(size = 10),
+    axis.text.y = element_text(size = 10),
+    axis.ticks.y = element_blank()
+  )
+
+```
+
+@fig-linePlot3 on the next page compares the sales count and redemption count in year 2021 and 2023
+
+1 emm这里写一点描述fig3的对比？，render后看页面排版
+
+```{r}
+#| label: fig-linePlot3
+#| fig-cap: Redemption and Sales Counts Over Time in 2021 and 2023
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-pos: "H"  #fix the plot in current place when render
+#| fig-width: 8
+#| fig-height: 4
+
+# Load necessary libraries
+library(ggplot2)
+library(dplyr)
+library(lubridate)
+library(gridExtra)  # To arrange plots
+
+# Filter the data for 2021 and 2023
+data_2023 <- data %>%
+  mutate(Timestamp = ymd_hms(Timestamp)) %>%
+  filter(year(Timestamp) == 2023)
+
+data_2021 <- data %>%
+  mutate(Timestamp = ymd_hms(Timestamp)) %>%
+  filter(year(Timestamp) == 2021)
+
+# Calculate the range for both y-axes for symmetry (using both years)
+y_max <- max(max(data_2021$Sales.Count), max(data_2021$Redemption.Count),
+             max(data_2023$Sales.Count), max(data_2023$Redemption.Count))
+
+# Plot for 2023
+plot_2023 <- ggplot(data_2023, aes(x = Timestamp)) +
+  geom_line(aes(y = Redemption.Count), color = "#A977A6", linewidth = 1.2) +
+  geom_line(aes(y = -Sales.Count), color = "#C1A9D2", linewidth = 1.2) +
+  scale_y_continuous(limits = c(-y_max, y_max), sec.axis = sec_axis(~ -., name = "Sales Count")) +
+  scale_x_datetime(date_breaks = "4 months", date_labels = "%b") +  # Show only 3 labels
+  labs(y = "Redemption Count", x = "Month", title = "2023") +
+  theme_minimal()
+
+# Plot for 2021
+plot_2021 <- ggplot(data_2021, aes(x = Timestamp)) +
+  geom_line(aes(y = Redemption.Count), color = "#F49600", linewidth = 1.2) +
+  geom_line(aes(y = -Sales.Count), color = "#F9CB80", linewidth = 1.2) +
+  scale_y_continuous(limits = c(-y_max, y_max), sec.axis = sec_axis(~ -., name = "Sales Count")) +
+  scale_x_datetime(date_breaks = "4 months", date_labels = "%b") +  # Show only 3 labels
+  labs(y = "Redemption Count", x = "Month", title = "2021") +
+  theme_minimal()
+
+# Arrange the two plots side by side
+grid.arrange(plot_2021, plot_2023, ncol = 2)
+
+```
+As shown in @fig-linePlot4, the x-axis denotes the timestamps of each record, while the y-axis represents the number of tickets processed.
+The blue line corresponds to the redemption counts, and the green line represents the sales counts, plotted inversely to emphasize the contrast between the two trends.
+The symmetrical y-axis allows for a clearer comparison between peaks and troughs in both counts over time.
+
+```{r}
+#| echo: false 
+#| warning: false
+#| message: false
+#| label: fig-linePlot4
+#| fig-cap: Redemption and Sales Counts Over Time in years
+#| fig-pos: "H"  #fix the plot in current place when render
+#| fig-width: 6
+#| fig-height: 2.5
 
@@
   # Line plot for Sales Count (inverted)
-  geom_line(aes(y = -Sales.Count), color = "#81B0B2", linewidth = 1.2) +
+  geom_line(aes(y = -Sales.Count), color = "#83A0BE", linewidth = 1.2) +
 
@@
     y = "Redemption Count",  # Y-axis for Redemption Count
-    x = "Timestamp"          # X-axis for Timestamp
+    x = "Years"          # X-axis for Timestamp
   ) +
@@
     # Color the left-side axis title (Redemption Count) to match the pink line
-    axis.title.y.left = element_text(color = "#81B0B2", size = 12),
+    axis.title.y.left = element_text(color = "#83A0BE", size = 12),
     # Center the title and increase font size
   )
+
 ```
 
-As shown in @fig-linePlot, the x-axis denotes the timestamps of each record, while the y-axis represents the number of tickets processed.
-The blue line corresponds to the redemption counts, and the green line represents the sales counts, plotted inversely to emphasize the contrast between the two trends.
-The symmetrical y-axis allows for a clearer comparison between peaks and troughs in both counts over time.
-
-The plot reveals significant variations in ferry usage over time.
+The plot reveals significant variations in ferry usage over years.
 We observe several peaks in both redemption and sales counts, likely aligning with periods of high activity, such as weekends, holidays, or the peak tourist season during the summer months.
@@
 This visualization allows us to understand whether there is a correlation between the number of tickets sold and the number of tickets redeemed during the same time period.
+
 ```{r}
@@
 #| message: false
+#| fig-width: 7
+#| fig-height: 3
 
@@
 #| message: false
+#| fig-width: 7
+#| fig-height: 2.5
 
@@
 
-In Section 2.2, the time series analysis of ticket sales and redemptions (Figure @fig-linePlot) revealed clear seasonal patterns, with pronounced peaks during the warmer months.
+In @sec-data-temporal, the time series analysis of ticket sales and redemptions (Figure @fig-linePlot1, @fig-linePlot2, @fig-linePlot3 and  @fig-linePlot4 ) revealed clear seasonal patterns, with pronounced peaks during the warmer months.
 This trend aligns with the historical and cultural significance of the Toronto Islands as a recreational destination, as documented by @varga1987toronto.

### 2024-09-27 outputs/paper/paper.tex
rendered the file to check the format of the paper

@@
 \hypersetup{
-  pdftitle={Analyzing Toronto Island Ferry Ticket Counts},
+  pdftitle={Uncovering Ferry Traffic Patterns: A Data-Driven Study of Toronto Island Ticket Sales and Redemptions},
   pdfauthor={Yingke He},
@@
 
-\title{Analyzing Toronto Island Ferry Ticket Counts\thanks{Code and data
-are available at: https://github.com/ohyykk/ferrySale.git}}
-\usepackage{etoolbox}
-\makeatletter
-\providecommand{\subtitle}[1]{% add subtitle to \maketitle
-  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
-}
-\makeatother
-\subtitle{A Statistical Exploration of Sales Trends}
+\title{Uncovering Ferry Traffic Patterns: A Data-Driven Study of Toronto
+Island Ticket Sales and Redemptions\thanks{Code and data are available
+at: https://github.com/ohyykk/ferrySale.git}}
 \author{Yingke He}
@@
 This study examines the ticket redemption and sales counts for the
-Toronto Island Ferry. We analyze time-based trends, relationships
-between sales and redemptions, and develop statistical models to predict
-ferry ticket usage. The results provide insights into usage patterns,
-highlight operational considerations, and suggest avenues for future
-research. Our analysis reveals distinct seasonal peaks and troughs in
-ferry usage, correlating with tourism and weather conditions.
-Furthermore, the study identifies discrepancies between ticket sales and
-redemptions, suggesting potential improvements in ticketing strategies
-and resource allocation.
+Toronto Island Ferry, analyzing time-based trends and the relationships
+between sales and redemptions. Statistical models are developed to
+predict ferry ticket usage. The results uncover seasonal trends,
+operational inefficiencies, and discrepancies between ticket sales and
+redemptions, highlighting the need for improved resource management and
+service adjustments to better align with fluctuating passenger demand.
+These findings offer insights into usage patterns, operational
+considerations, and suggest avenues for future research, particularly in
+optimizing ticketing strategies and resource allocation.
 \end{abstract}
@@
 biodiversity and recreation. Additionally, the potential role of ferries
-in urban transit systems, as explored by Kopystynski (1979), underscores
+in urban transit systems, as explained by Kopystynski (1979), shows
 their importance in enhancing urban mobility and connectivity.
 
-This paper aims to analyze the ticket counts for the Toronto Island
-Ferry, exploring trends in ticket redemption and sales, and drawing
-connections between historical developments and current usage patterns.
-By integrating insights from both historical and modern perspectives,
-including contemporary studies on ferry commuting from Roseman (2020),
-this analysis seeks to provide a comprehensive understanding of the
-factors influencing ferry usage. The remainder of this paper is
-structured as follows: Section 2 describes the dataset used for
-analysis, Section 3 outlines the modeling approach, Section 4 presents
-the results, and Section 5 discusses the implications of the findings
-and potential directions for future research. We use R Core Team (2023)
-and Wickham et al. (2019a), and We use Section~\ref{sec-data} to
-reference the data section.
+With the ticket counts for the Toronto Island Ferry and the trends in
+ticket redemption and sales, we can find the connections between
+historical developments and current usage patterns. By combining both
+historical and modern perspectives and including contemporary studies on
+ferry commuting from Roseman (2020), the analysis provides a
+comprehensive understanding of the factors influencing ferry usage.
+Statistical models are used to analyze ferry ticket demand while showing
+the seasonal fluctuations and operational challenges.
+
+Despite the longstanding role of the Toronto Island Ferry system in
+connecting downtown Toronto with the Toronto Islands, there is limited
+research on the detailed patterns of ferry usage and the efficiency of
+the service. While the ferry system serves both residents and tourists,
+the lack of data-driven analysis makes it harder to optimize ferry
+operations, manage peak demand periods, and improve ticketing
+strategies. It is important to fill this gap because of the ferry's dual
+function of accommodating daily commuters alongside a significant influx
+of seasonal visitors, which introduces variability in service demand.
+Without a comprehensive understanding of the underlying trends in ticket
+sales and redemptions, it becomes challenging to address inefficiencies
+or anticipate future operational challenges. We will address this
+problem by providing a detailed analysis of ticketing patterns and
+offering insights into how the ferry system can better meet passenger
+needs through data-informed decision-making.
+
+In addition to the expected peaks during high tourism periods, this
+study investigates inconsistencies between tickets sold and redeemed,
+offering insights into potential service inefficiencies. The dual role
+of the Toronto Island Ferry system in accommodating both daily commuters
+and a large number of seasonal visitors makes it an important case study
+in urban transportation planning.
+
+The remainder of this paper is structured as follows:
+Section~\ref{sec-data} introduces the dataset used for the analysis,
+explaining the key variables and presenting key data analysis.
+Section~\ref{sec-discussion} discusses the implications of the findings
+and potential directions for future research. Additional information is
+presented in the appendix at Section~\ref{sec-appendix}. R Core Team
+(2023) and Wickham et al. (2019a) is utilized for the coding proposes of
+this paper, and the data section is referenced using
+Section~\ref{sec-data}.
 
@@
 
+\subsection{Measurements}\label{measurements}
+
+This report utilizes the Toronto Island Ferry Ticket Counts dataset
+published by the City of Toronto on the Open Data Portal(City of Toronto
+(2024)). Specifically, time-stamped records of both ticket sales and
+redemptions were recoreded. Every time a ticket is sold or redeemed, it
+represents a physical action by a ferry passenger in Toronto traveling
+to the Toronto Island. Actions are then recorded by the ticketing
+systems, which automatically records the transaction timestamp, along
+with the number of tickets sold and redeemed at that specific time.
+
+This process converts individual passenger transactions into structured
+dataset entries, with timestamps to enable tracking over time. The data
+captures real-world trends such as higher ferry activity during peak
+tourism seasons or weekends, providing useful insights into ferry usage
+patterns. It enables further research to analyze passenger demand and
+operational trends, helping to make more effective ferry management
+decisions.
+
 \subsection{Results}\label{sec-data-resultsk}
 
-The table Table~\ref{tbl-cleaned_data} provides a detailed snapshot of
-the Toronto Island Ferry ticket transactions for a specific time period.
-It includes a unique ID for each transaction, the Time at which each
-transaction occurred, the number of Tickets Redeemed, the number of
-Tickets Sold, and the Date of the transaction.
+Table~\ref{tbl-cleaned_data} provides a detailed snapshot of the Toronto
+Island Ferry ticket transactions for a specific time period. It includes
+a unique ID for each transaction, the Time at which each transaction
+occurred, the number of Tickets Redeemed, the number of Tickets Sold,
+and the Date of the transaction.
 
@@
 
-Sample Cleaned Data
+The above table shows the typical flucuations in ticket sales and
+redemptions throughout the day. For example, during late evening hours,
+we observe varying ticket redemption and sales patterns, which may
+correspond to changes in passenger demand as the ferry service
+approaches the end of its operational hours.
 
-This table shows the typical flucuations in ticket sales and redemptions
-throughout the day. For example, during late evening hours, we observe
-varying ticket redemption and sales patterns, which may correspond to
-changes in passenger demand as the ferry service approaches the end of
-its operational hours.
+\begin{longtable}[]{@{}ccccc@{}}
+\caption{Summary Statistics of Ticket
+Counts}\label{tbl-summary_stats}\tabularnewline
+\toprule\noalign{}
+Avg Redeemed & Avg Sold & Max Redeemed & Max Sold & Count \\
+\midrule\noalign{}
+\endfirsthead
+\toprule\noalign{}
+Avg Redeemed & Avg Sold & Max Redeemed & Max Sold & Count \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+48.44548 & 49.36013 & 7216 & 7229 & 229380 \\
+\end{longtable}
 
-\subsection{Redemption and Sales Counts Over
-Time}\label{redemption-and-sales-counts-over-time}
+The summary statistics, shown in Table~\ref{tbl-summary_stats}, provide
+a clear understanding of ferry ticket trends. The high volume of tickets
+sold and redeemed indicates strong demand for the ferry service, with
+the close alignment between sales and redemptions suggesting high
+customer engagement. On average, 46.36 tickets were sold every 15
+minutes during working hours, and 48.45 tickets were redeemed,
+highlighting that most customers have used their tickets. The small gap
+between sales and redemptions implies a few no-shows or unused tickets.
+The maximum values of 7,216 for sales and 7,229 for redemptions
+demonstrate the ferry's capacity to handle high passenger volumes,
+typically operating at full capacity on busy days.
+
+\subsubsection{Temporal Trends in Redemption and Sales
+Counts}\label{sec-data-temporal}
 
@@
 ticket redemption and sales counts over time. The line plot shown in
-Figure~\ref{fig-linePlot} provides a clear visual representation of
-these trends, serving as a starting point for deeper analysis of ferry
-usage patterns.
+Figure~\ref{fig-linePlot1}, Figure~\ref{fig-linePlot2},
+Figure~\ref{fig-linePlot3} and Figure~\ref{fig-linePlot4} provides a
+clear visual representation of these trends, serving as a starting point
+for deeper analysis of ferry usage patterns.
 
-The dataset used in this analysis includes timestamped records of ticket
-redemptions and sales for the Toronto Island Ferry system, offering a
-time series perspective on ferry usage. The plot in
-Figure~\ref{fig-linePlot} specifically highlights how these counts
-fluctuate over time, reflecting the flow of passenger demand. The y-axis
-for redemption counts is displayed in blue, while the y-axis for sales
-counts is displayed in green, but inverted to better show their
-correlation.
+The peak day analysis table below highlights the highest ferry ticket
+sales, often aligning with public holidays or major events, such as Aug
+1 2016, and July 1 2018. On September 15, 2020, ferry sales saw a
+significant drop, with only 5,003 tickets sold. This decline could be
+due to the impact of COVID-19(Lee and Leung (2021)).
+
+\begin{longtable}[]{@{}rlr@{}}
+\caption{Table: Peak Day for Each Year Based on Total
+Sales}\tabularnewline
+\toprule\noalign{}
+Year & Peak Day & Total Sales \\
+\midrule\noalign{}
+\endfirsthead
+\toprule\noalign{}
+Year & Peak Day & Total Sales \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+2015 & 2015-07-18 & 23354 \\
+2016 & 2016-08-01 & 25552 \\
+2017 & 2017-08-20 & 17322 \\
+2018 & 2018-07-01 & 23669 \\
+2019 & 2019-08-04 & 20911 \\
+2020 & 2020-09-05 & 5003 \\
+2021 & 2021-09-05 & 12356 \\
+2022 & 2022-07-16 & 20419 \\
+2023 & 2023-08-19 & 20203 \\
+2024 & 2024-07-13 & 19603 \\
+\end{longtable}
+
+The dataset used in this analysis provides timestamped records of ticket
+redemptions and sales for the Toronto Island Ferry system, allowing for
+a detailed time series analysis of ferry usage patterns. As illustrated
+in Figure~\ref{fig-linePlot1}, ferry usage fluctuates significantly
+throughout the year, with peak activity consistently observed between
+July and September. This pattern corresponds to the increased passenger
+demand during the summer holiday season and major events. In 2020,
+ticket demand saw a sharp decline, likely due to the impact of the
+COVID-19 pandemic(Urbanyi-Popiolek (2020)), but it showed signs of
+gradual recovery starting in 2021 as restrictions eased and tourism
+began to rebound in Toronto(Government of Canada (2024)). By identifying
+these peak traffic periods in advance, ferry operators can implement
+targeted operational strategies, such as scheduling additional ferry
+trips or introducing crowd control measures to minimize the risk of
+overcrowding or service disruptions. Identifing these high-demand
+periods can ensure smoother ferry operations and enhance the overall
+passenger experience.
 
@@
 
-\includegraphics{paper_files/figure-pdf/fig-linePlot-1.pdf}
+\includegraphics{paper_files/figure-pdf/fig-linePlot1-1.pdf}
 
@@
 
-\caption{\label{fig-linePlot}Redemption and Sales Counts Over Time}
+\caption{\label{fig-linePlot1}Redemption and Sales Counts Over Time in
+days}
 
@@
 
-As shown in Figure~\ref{fig-linePlot}, the x-axis denotes the timestamps
-of each record, while the y-axis represents the number of tickets
-processed. The blue line corresponds to the redemption counts, and the
-green line represents the sales counts, plotted inversely to emphasize
-the contrast between the two trends. The symmetrical y-axis allows for a
-clearer comparison between peaks and troughs in both counts over time.
+加一小段描述描述figure1
 
-The plot reveals significant variations in ferry usage over time. We
+Figure~\ref{fig-linePlot2} below shows the Redemption and Sales Counts
+in months over years， 加一段描述figure2
+
+\newpage
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics{paper_files/figure-pdf/fig-linePlot2-1.pdf}
+
+}
+
+\caption{\label{fig-linePlot2}Redemption and Sales Counts Over Time in
+months}
+
+\end{figure}%
+
+Figure~\ref{fig-linePlot3} on the next page compares the sales count and
+redemption count in year 2021 and 2023
+
+1 emm这里写一点描述fig3的对比？，render后看页面排版
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics{paper_files/figure-pdf/fig-linePlot3-1.pdf}
+
+}
+
+\caption{\label{fig-linePlot3}Redemption and Sales Counts Over Time in
+2021 and 2023}
+
+\end{figure}%
+
+As shown in Figure~\ref{fig-linePlot4}, the x-axis denotes the
+timestamps of each record, while the y-axis represents the number of
+tickets processed. The blue line corresponds to the redemption counts,
+and the green line represents the sales counts, plotted inversely to
+emphasize the contrast between the two trends. The symmetrical y-axis
+allows for a clearer comparison between peaks and troughs in both counts
+over time.
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics{paper_files/figure-pdf/fig-linePlot4-1.pdf}
+
+}
+
+\caption{\label{fig-linePlot4}Redemption and Sales Counts Over Time in
+years}
+
+\end{figure}%
+
+The plot reveals significant variations in ferry usage over years. We
 observe several peaks in both redemption and sales counts, likely
@@
 
-\subsection{Redemption vs Sales Count}\label{redemption-vs-sales-count}
+\subsubsection{Comparative Analysis of Ticket Sales and
+Redemptions}\label{comparative-analysis-of-ticket-sales-and-redemptions}
 
@@
 
-\newpage
-
-\subsection{Side by Side Comparison}\label{side-by-side-comparison}
+\subsubsection{Visual Comparison of Ticket Sales and Redemption
+Distribution}\label{visual-comparison-of-ticket-sales-and-redemption-distribution}
 
@@
 
-In Section 2.2, the time series analysis of ticket sales and redemptions
-(Figure Figure~\ref{fig-linePlot}) revealed clear seasonal patterns,
-with pronounced peaks during the warmer months. This trend aligns with
-the historical and cultural significance of the Toronto Islands as a
+In Section~\ref{sec-data-temporal}, the time series analysis of ticket
+sales and redemptions (Figure Figure~\ref{fig-linePlot1},
+Figure~\ref{fig-linePlot2}, Figure~\ref{fig-linePlot3} and
+Figure~\ref{fig-linePlot4} ) revealed clear seasonal patterns, with
+pronounced peaks during the warmer months. This trend aligns with the
+historical and cultural significance of the Toronto Islands as a
 recreational destination, as documented by Varga (1987). The increased
@@
 
-The scatter plot analysis in Section 2.3 (Figure
-Figure~\ref{fig-scatterPlot}) demonstrated a generally positive
-correlation between ticket sales and redemptions, suggesting that most
-tickets purchased are used. However, discrepancies where sales exceed
-redemptions could be linked to bulk purchases, advanced ticket sales, or
-unredeemed tickets due to factors like adverse weather conditions. This
-observation is consistent with the analysis by Kopystynski (1979), who
-highlighted the complexities of integrating ferry systems into urban
-transit networks. Understanding these discrepancies is critical for
-improving ticketing policies and service offerings to better match
-supply with demand.
+The scatter plot analysis in Section 2.3 (Figure~\ref{fig-scatterPlot})
+demonstrated a generally positive correlation between ticket sales and
+redemptions, suggesting that most tickets purchased are used. However,
+discrepancies where sales exceed redemptions could be linked to bulk
+purchases, advanced ticket sales, or unredeemed tickets due to factors
+like adverse weather conditions. This observation is consistent with the
+analysis by Kopystynski (1979), who highlighted the complexities of
+integrating ferry systems into urban transit networks. Understanding
+these discrepancies is critical for improving ticketing policies and
+service offerings to better match supply with demand.
 
@@
 \begin{CSLReferences}{1}{0}
+\bibitem[\citeproctext]{ref-toronto_ferry_ticket_counts_2024}
+City of Toronto. 2024. {``Toronto Island Ferry Ticket Counts.''}
+\url{https://open.toronto.ca/dataset/toronto-island-ferry-ticket-counts/}.
+
+\bibitem[\citeproctext]{ref-travel_gov_2024}
+Government of Canada. 2024. {``Travel and COVID-19.''}
+\url{https://travel.gc.ca/travel-covid}.
+
 \bibitem[\citeproctext]{ref-kopystynski1979potential}
@@
 Columbia.
+
+\bibitem[\citeproctext]{ref-lee2021island}
+Lee, H. Y., and K. Y. K. Leung. 2021. {``Island Ferry Travel During
+COVID-19: Charting the Recovery of Local Tourism in Hong Kong.''}
+\emph{Current Issues in Tourism} 25 (1): 76--93.
+\url{https://doi.org/10.1080/13683500.2021.1911964}.
 
@@
 
+\bibitem[\citeproctext]{ref-urbanyi2020maritime}
+Urbanyi-Popiolek, Ilona. 2020. {``Maritime Tourism in the Time of
+COVID-19 Pandemic in the Baltic Sea Region--Challenges for Ferry and
+Cruise Operators.''} In \emph{Economic and Social Development: Book of
+Proceedings}, 397--405.
+
 \bibitem[\citeproctext]{ref-varga1987toronto}
@@
 
+\bibitem[\citeproctext]{ref-tidyverse2019}
+Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy
+D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019b.
+{``Welcome to the Tidyverse.''} \emph{Journal of Open Source Software} 4
+(43): 1686. \url{https://doi.org/10.21105/joss.01686}.
+
 \bibitem[\citeproctext]{ref-rohan}
@@
 4 (43): 1686. \url{https://doi.org/10.21105/joss.01686}.
-
-\bibitem[\citeproctext]{ref-wickham2019tidyverse}
-Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy
-D'Agostino McGowan, Romuald François, Garrett Grolemund, et al. 2019b.
-{``Welcome to the Tidyverse.''} \emph{Journal of Open Source Software}.
-\url{https://doi.org/10.21105/joss.01686}.
 

### 2024-09-27 outputs/paper/paper.qmd
made a few amendments on the main text
add a few details on data analsis
add more informations on the data cleaning part in appendix section

@@
 
-The dataset used in this analysis is the 2024 installment of “Toronto Island Ferry Ticket Counts” obtained from OpenDataToronto (@varga1987toronto).
+The dataset used in this analysis is the 2024 installment of “Toronto Island Ferry Ticket Counts” obtained from OpenDataToronto (@R-opendatatoronto).
 This dataset contains information about the number of ferry tickets redeemed and sold for the Toronto Island Ferry system, which serves as a critical transportation link between downtown Toronto and the Toronto Islands.
@@
 The data was processed and analyzed using the R programming language (@citeR).
-The tidyverse (@tidyverse2019) and opendatatoronto packages were employed to download, clean, and prepare the dataset for analysis.
+The tidyverse (@citeTidyverse) and opendatatoronto packages were employed to download, clean, and prepare the dataset for analysis.
 The resulting cleaned dataset was then utilized to generate the visualizations and statistical analyses presented in this paper using a collection of libraries which included the following packages:
 
-- ggplot2 [@wickham2016ggplot2]
-- knitr [@xie2023knitr]
-- readr [@wickham2023readr]
+- ggplot2 [@citeGgplot2]
+- knitr [@citeKnitr]
+- readr [@citeReadr]
+- gridExtra, solely for the propose of plot arrangements [@R-gridExtra]
+- lubridate [@R-lubridate]
+- here, helps to create relative file paths [@citeHere]
 
@@
 The dataset used in this analysis provides timestamped records of ticket redemptions and sales for the Toronto Island Ferry system, allowing for a detailed time series analysis of ferry usage patterns.
-As illustrated in @fig-linePlot1, ferry usage fluctuates significantly throughout the year, with peak activity consistently observed between July and September.
-This pattern corresponds to the increased passenger demand during the summer holiday season and major events.
-In 2020, ticket demand saw a sharp decline, likely due to the impact of the COVID-19 pandemic(@urbanyi2020maritime), but it showed signs of gradual recovery starting in 2021 as restrictions eased and tourism began to rebound in Toronto(@travel_gov_2024).
-By identifying these peak traffic periods in advance, ferry operators can implement targeted operational strategies, such as scheduling additional ferry trips or introducing crowd control measures to minimize the risk of overcrowding or service disruptions.
-Identifing these high-demand periods can ensure smoother ferry operations and enhance the overall passenger experience.
+@fig-linePlot1 illustrates the peak day of ticket sales for each year.
+Understanding these peak days and their contributing factors can help in forecasting future demand and optimizing resource allocation for similar high-traffic periods.
 
@@
 #| label: fig-linePlot1
-#| fig-cap: Redemption and Sales Counts Over Time in days
+#| fig-cap: Peak Day for Each Year
 #| echo: false
@@
 #| fig-height: 4
+#| fig-pos: "H"  #fix the plot in current place when render
 #| 
@@
 
-@fig-linePlot2 below shows the Redemption and Sales Counts in months over years 加一段描述plot2
+@fig-linePlot1 presents the daily peak ticket sales counts for the Toronto Island Ferry system across various years, offering a clear visualization of seasonal fluctuations in ferry usage.
+Each line represents the total ticket sales for the busiest day of each month throughout the year, with distinct colors assigned to different years for easy comparison.
+Notably, the graph highlights consistent peaks in ticket sales during the summer months, especially from June to August, reflecting heightened ferry usage during the tourist season.
+The sharp decline in sales in 2020, evident in the orange line, corresponds to the COVID-19 pandemic's impact on public transportation and tourism activities (@urbanyi2020maritime).
+As restrictions eased in 2021, a gradual recovery in ticket sales is observed, depicted by the brown line, indicating a resurgence in ferry demand as tourism began to pick up in Toronto (@travel_gov_2024).
+This visualization not only showcases the annual variations in peak demand but also underscores the need for adaptive operational strategies to accommodate seasonal surges and unexpected disruptions.
 
-\newpage
+@fig-linePlot2 on the following page shows the Redemption and Sales Counts in months over years.
+The graph uses dashed lines to represent the monthly sales counts and straight lines for the redemption counts.
+Notably, the inverted redemption counts are plotted for better visualization alongside the sales data.
+This figure highlights consistent seasonal patterns, with both ticket sales and redemptions peaking during the summer months, from June to August, across most years.
+The dip observed in 2020 reflects the significant impact of the COVID-19 pandemic on ferry operations, with reduced ferry usage compared to previous years.
+These trends show the importance of flexible operational planning to accommodate both predictable seasonal demand and unexpected disruptions.
 
@@
 #| echo: false
-#| fig-pos: "H"  #fix the plot in current place when render
 #| fig-width: 8
-#| fig-height: 4
+#| fig-height: 3.5
 #| 
@@
 
-@fig-linePlot3 on the next page compares the sales count and redemption count in year 2021 and 2023
-
-1 emm这里写一点描述fig3的对比？，render后看页面排版
+@fig-linePlot3 below compares the sales count and redemption count in year 2021 and 2023. The two side-by-side plots illustrate distinct usage patterns across these years.
 
@@
 #| message: false
-#| fig-pos: "H"  #fix the plot in current place when render
+#| fig-pos: "H"  # Fix the plot in current place when rendered
 #| fig-width: 8
-#| fig-height: 4
+#| fig-height: 3
 
@@
 
-# Plot for 2023
+# Plot for 2023 (Hide negative values)
 plot_2023 <- ggplot(data_2023, aes(x = Timestamp)) +
@@
   geom_line(aes(y = -Sales.Count), color = "#C1A9D2", linewidth = 1.2) +
-  scale_y_continuous(limits = c(-y_max, y_max), sec.axis = sec_axis(~ -., name = "Sales Count")) +
+  scale_y_continuous(
+    limits = c(-y_max, y_max), 
+    breaks = seq(0, y_max, by = 1000),  # Only show positive labels for Redemption Count
+    sec.axis = sec_axis(~ -., name = "Sales Count", breaks = seq(0, y_max, by = 1000))  # Only show positive labels for Sales Count
+  ) +
   scale_x_datetime(date_breaks = "4 months", date_labels = "%b") +  # Show only 3 labels
   labs(y = "Redemption Count", x = "Month", title = "2023") +
-  theme_minimal()
+  theme_minimal() +
+  theme(
+    axis.title.y.right = element_text(color = "#C1A9D2"),
+    axis.title.y.left = element_text(color = "#A977A6")
+  )
 
-# Plot for 2021
+# Plot for 2021 (Hide negative values)
 plot_2021 <- ggplot(data_2021, aes(x = Timestamp)) +
@@
   geom_line(aes(y = -Sales.Count), color = "#F9CB80", linewidth = 1.2) +
-  scale_y_continuous(limits = c(-y_max, y_max), sec.axis = sec_axis(~ -., name = "Sales Count")) +
+  scale_y_continuous(
+    limits = c(-y_max, y_max), 
+    breaks = seq(0, y_max, by = 1000),  # Only show positive labels for Redemption Count
+    sec.axis = sec_axis(~ -., name = "Sales Count", breaks = seq(0, y_max, by = 1000))  # Only show positive labels for Sales Count
+  ) +
   scale_x_datetime(date_breaks = "4 months", date_labels = "%b") +  # Show only 3 labels
   labs(y = "Redemption Count", x = "Month", title = "2021") +
-  theme_minimal()
+  theme_minimal() +
+  theme(
+    axis.title.y.right = element_text(color = "#F9CB80"),
+    axis.title.y.left = element_text(color = "#F49600")
+  )
 
-# Arrange the two plots side by side
+# Corrected function call
 grid.arrange(plot_2021, plot_2023, ncol = 2)
@@
 ```
+
+In 2021, ticket sales and redemptions were more variable, reflecting the gradual recovery from the COVID-19 pandemic as restrictions were lifted and ferry services resumed normal operations.
+Peaks and troughs are more pronounced, indicating fluctuating demand as public confidence in using public transportation returned.
+Conversely, the patterns in 2023 exhibit a more consistent trend, with a steadier increase in both ticket sales and redemptions during the summer months, reflecting a return to pre-pandemic demand levels.
+This comparison highlights the evolving dynamics in ferry usage, influenced by external factors such as public health measures and tourism trends.
+
 As shown in @fig-linePlot4, the x-axis denotes the timestamps of each record, while the y-axis represents the number of tickets processed.
-The blue line corresponds to the redemption counts, and the green line represents the sales counts, plotted inversely to emphasize the contrast between the two trends.
+The red line corresponds to the redemption counts, and the blue line represents the sales counts, plotted inversely to emphasize the contrast between the two trends.
 The symmetrical y-axis allows for a clearer comparison between peaks and troughs in both counts over time.
@@
 ```{r}
+#| label: fig-linePlot4
+#| fig-cap: Redemption and Sales Counts Over Time in years
+#| fig-pos: "H"
+#| fig-width: 6
+#| fig-height: 2.5
 #| echo: false
@@
 #| message: false
-#| label: fig-linePlot4
-#| fig-cap: Redemption and Sales Counts Over Time in years
-#| fig-pos: "H"  #fix the plot in current place when render
-#| fig-width: 6
-#| fig-height: 2.5
 
@@
   # Line plot for Redemption Count
-  geom_line(aes(y = Redemption.Count), color = "#F7C4C1", linewidth = 1.2) +
+  geom_line(aes(y = Redemption.Count), color = "brown2", linewidth = 1.2) +
   # Line plot for Sales Count (inverted)
-  geom_line(aes(y = -Sales.Count), color = "#83A0BE", linewidth = 1.2) +
+  geom_line(aes(y = -Sales.Count), color = "#377EB8", linewidth = 1.2) +
 
   # Set y-axis limits symmetrically and apply the secondary axis
-  scale_y_continuous(limits = c(-y_max, y_max),
-                     breaks = seq(-y_max, y_max, by = 1000),
+  scale_y_continuous(
+    limits = c(-y_max, y_max),
+    breaks = seq(0, y_max, by = 1000),  # Show only positive values for Redemption Count
     sec.axis = sec_axis(~ -., 
                         name = "Sales Count",
-                                         breaks = seq(y_max, -y_max, by = -1000))) +
+                        breaks = seq(0, y_max, by = 1000))  # Show only positive values for Sales Count
+  ) +
 
@@
   theme(
-    # Color the right-side axis title (Sales Count) to match the orange line
-    axis.title.y.right = element_text(color = "#F7C4C1", size = 12),
-    # Color the left-side axis title (Redemption Count) to match the pink line
-    axis.title.y.left = element_text(color = "#83A0BE", size = 12),
+    # Color the right-side axis title (Sales Count) to match the Sales line
+    axis.title.y.left = element_text(color = "brown3", size = 12),
+    # Color the left-side axis title (Redemption Count) to match the Redemption line
+    axis.title.y.right = element_text(color = "dodgerblue3", size = 12),
     # Center the title and increase font size
@@
 
+While the analysis provides valuable insights, there are some limitations that should be acknowledged.
+The dataset does not account for external variables such as weather conditions or special events, which may also significantly impact ferry usage.
+Additionally, the data is limited to ticket counts and does not capture other relevant factors such as passenger demographics or satisfaction levels.
+
 \newpage
@@
 
-The data cleaning process involved filtering out some of the columns from the raw dataset and renaming some of the data entries for clarity and simplicity.
+The raw FerrySale dataset underwent several cleaning steps to ensure its accuracy and usability for analysis.
+Initially, all rows with missing values were removed, followed by converting the Timestamp column to a standardized date-time format for consistent time series analysis.
+Duplicate entries were identified and eliminated to prevent inflated ticket counts.
+Additionally, transactions with unrealistic values, such as negative counts for ticket redemptions or sales, were filtered out.
+The resulting cleaned dataset, saved as cleaned_data.csv, provided a reliable foundation for the statistical analyses and visualizations presented in this paper.
 

### 2024-09-27 outputs/paper/paper.qmd
Final draft of the paper: checked and made minor changes to the main text

@@
 ## Results {#sec-data-resultsk}
+
+The raw data underwent extensive cleaning, including the removal of rows with missing values, conversion of timestamps to a uniform format, and filtering out duplicates and unrealistic values such as negative ticket counts.
+This process ensured a reliable foundation for subsequent analysis.
 
@@
 #| fig-width: 8
-#| fig-height: 3.5
+#| fig-height: 3.3
 #| 
@@
                      breaks = seq(-max_count, max_count, by = 100000),
-                     sec.axis = sec_axis(~ -., name = "Sales Count (dashed lines)")) +
+                     sec.axis = sec_axis(~ -., name = "Sales Count (straight lines)")) +
 
@@
   labs(
-    y = "Redemption Count (straight lines)",  # Y-axis for Redemption Count (inverted)
+    y = "Redemption Count (dashed lines)",  # Y-axis for Redemption Count (inverted)
     x = "Month",             # X-axis for Month
@@
 
-@fig-linePlot3 below compares the sales count and redemption count in year 2021 and 2023. The two side-by-side plots illustrate distinct usage patterns across these years.
-
 ```{r}
@@
 #| message: false
-#| fig-pos: "H"  # Fix the plot in current place when rendered
 #| fig-width: 8
-#| fig-height: 3
+#| fig-height: 2.8
 
@@
 ```
+@fig-linePlot3 compares the sales count and redemption count in year 2021 and 2023. The two side-by-side plots illustrate distinct usage patterns across these years.
 

### 2024-09-27 outputs/paper/paper.qmd
final draft upoaded

@@
 
-@fig-linePlot2 on the following page shows the Redemption and Sales Counts in months over years.
+@fig-linePlot2 shows the Redemption and Sales Counts in months over years.
 The graph uses dashed lines to represent the monthly sales counts and straight lines for the redemption counts.
@@
   # Set symmetrical y-axis limits for both Sales and Redemption
-  scale_y_continuous(limits = c(-max_count, max_count),
+  scale_y_continuous(
+    limits = c(-max_count, max_count),
     breaks = seq(-max_count, max_count, by = 100000),
-                     sec.axis = sec_axis(~ -., name = "Sales Count (straight lines)")) +
+    labels = function(x) abs(x),  # Remove the negative signs from lower half labels
+    sec.axis = sec_axis(~ -., name = "Sales Count (straight lines)")
+  ) +
 
@@
     axis.title.y.left = element_text(color = "black", size = 12, margin = margin(l = 10)),
+    axis.text.y.right = element_blank(),  # Hide axis values of the right y-axis
     plot.title = element_text(hjust = 0.5, size = 16),

### 2024-09-27 outputs/paper/paper.qmd
made a very minor change
change from sales to Sales

@@
 @fig-linePlot2 shows the Redemption and Sales Counts in months over years.
-The graph uses dashed lines to represent the monthly sales counts and straight lines for the redemption counts.
-Notably, the inverted redemption counts are plotted for better visualization alongside the sales data.
+The graph uses straight lines to represent the monthly sales counts and dashed lines for the redemption counts.
+Notably, the inverted sales counts are plotted for better visualization alongside the redemption data.
 This figure highlights consistent seasonal patterns, with both ticket sales and redemptions peaking during the summer months, from June to August, across most years.
